---
title: "Exclusion criteria"
author: "Lori,Mitchell, Rachel K.Y. Tsui, & Krista Byers-Heinlein"
date: '`r format(Sys.time(), "%a %b %d %X %Y")`'
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: yes
---

```{r setup, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(tidylog)
library(matrixStats)
```

# Intro
This script documents how we screened the demographic data from the webCDI data and identified those participants to be excluded.

# Read in data
```{r cache=TRUE}
clean_demog <- read.csv(here::here("data_keepers/public_clean_demog.csv"))
```

# Participant info (total)
```{r cache=TRUE}
# Number of total participants (after removing incompleted entries, monolinguals, and pilot data)
clean_demog %>%
  filter(monolingual != 1) %>%
  distinct(subject_id) %>%
  nrow()

# Number of girls and boys
clean_demog %>%
  filter(monolingual != 1) %>%
  select(subject_id, sex) %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  dplyr::count(sex)
```

# Participant info after keeping those where both English and French CDIs were completed
```{r cache=TRUE}
# Number of total participants (after removing incompleted entries, monolinguals, and pilot data)
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  distinct(subject_id) %>%
  nrow()

# Number of girls and boys
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, sex) %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  dplyr::count(sex)

# Number of participants who contributed data more than one time point
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  group_by(subject_id) %>%
  dplyr::count(subject_id) %>%
  filter(n > 1) %>%
  nrow()

# Age range (across all participants and all months)
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  summarize(mean_age = mean(age_months, na.rm=T),
            sd_age = sd(age_months, na.rm=T),
            min_age = min(age_months, na.rm=T),
            max_age = max(age_months, na.rm=T))

# Age range for 1st month data entry
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  filter(n_months == 1) %>%
  summarize(mean_age = mean(age_months, na.rm=T),
            sd_age = sd(age_months, na.rm=T),
            min_age = min(age_months, na.rm=T),
            max_age = max(age_months, na.rm=T))

# Age range for final month data entry
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  arrange(subject_id, n_months) %>%
  # get the last row of data within each participant
  group_by(subject_id) %>%
  slice(n()) %>%
  ungroup() %>%
  # summarize
  summarize(mean_age = mean(age_months, na.rm=T),
            sd_age = sd(age_months, na.rm=T),
            min_age = min(age_months, na.rm=T),
            max_age = max(age_months, na.rm=T)) # the oldest age is 27m1
  
```

# Average language exposure (weighted mean is calculated as participants contributed different number of LEQ update data)
```{r cache=TRUE}
# by language
## minimum and maximum of individual averages
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant
  summarize(mean_exposure_eng = mean(exposure_eng),
            mean_exposure_fr = mean(exposure_fr),
            mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  # minimum and maximum of the individual averages
  pivot_longer(c(mean_exposure_eng, mean_exposure_fr, mean_exposure_other), names_to = "language", values_to = "mean_exposure") %>%
  group_by(language) %>%
  summarize(min_exposure = min(mean_exposure),
            max_exposure = max(mean_exposure))

## weighted mean across all participants
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  mutate(count = n()) %>%
  ungroup() %>%
  pivot_longer(c(exposure_eng, exposure_fr, exposure_other), names_to = "language", values_to = "exposure") %>%
  group_by(language) %>%
  summarize(mean_exposure = round(weightedMean(exposure, count), 1),
            sd_exposure = round(weightedSd(exposure, count), 1))


# by dominance
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  mutate(exposure_dom = if_else(exposure_eng > exposure_fr, exposure_eng, exposure_fr),
         exposure_nondom = if_else(exposure_eng > exposure_fr, exposure_fr, exposure_eng)) %>%
  group_by(subject_id) %>%
  mutate(count = n()) %>%
  ungroup() %>%
  pivot_longer(-c(subject_id:exposure_fr, count), names_to = "language", values_to = "exposure") %>%
  group_by(language) %>%
  summarize(mean_exposure = round(weightedMean(exposure, count), 1),
            sd_exposure = round(weightedSd(exposure, count), 1))


# Number of English dominant
## minimum and maximum of individual averages
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant
  summarize(mean_exposure_eng = mean(exposure_eng),
            mean_exposure_fr = mean(exposure_fr),
            mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  # English exposure > French exposure
  filter(mean_exposure_eng > mean_exposure_fr) %>%
  # minimum and maximum of the individual averages
  pivot_longer(c(mean_exposure_eng, mean_exposure_fr, mean_exposure_other), names_to = "language", values_to = "mean_exposure") %>%
  group_by(language) %>%
  summarize(number = length(unique(subject_id)),
            min_exposure = min(mean_exposure),
            max_exposure = max(mean_exposure))

## weighted mean across all participants
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant to define dominance
  mutate(count = n(),
         mean_exposure_eng = mean(exposure_eng),
         mean_exposure_fr = mean(exposure_fr),
         mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  filter(mean_exposure_eng > mean_exposure_fr) %>% # English exposure > French exposure
  select(-c(matches("mean"))) %>%
  # calculating weigthed mean
  pivot_longer(c(exposure_eng, exposure_fr, exposure_other), names_to = "language", values_to = "exposure") %>%
  group_by(language) %>%
  summarize(number = length(unique(subject_id)),
            mean_exposure = round(weightedMean(exposure, count), 1),
            sd_exposure = round(weightedSd(exposure, count), 1))


# Number of French dominant
## minimum and maximum of individual averages
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant
  summarize(mean_exposure_eng = mean(exposure_eng),
            mean_exposure_fr = mean(exposure_fr),
            mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  # English exposure < French exposure
  filter(mean_exposure_eng < mean_exposure_fr) %>%
  # minimum and maximum of the individual averages
  pivot_longer(c(mean_exposure_eng, mean_exposure_fr, mean_exposure_other), names_to = "language", values_to = "mean_exposure") %>%
  group_by(language) %>%
  summarize(number = length(unique(subject_id)),
            min_exposure = min(mean_exposure),
            max_exposure = max(mean_exposure))

## weighted mean across all participants
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant to define dominance
  mutate(count = n(),
         mean_exposure_eng = mean(exposure_eng),
         mean_exposure_fr = mean(exposure_fr),
         mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  filter(mean_exposure_eng < mean_exposure_fr) %>% # English exposure < French exposure
  select(-c(matches("mean"))) %>%
  # calculating weigthed mean
  pivot_longer(c(exposure_eng, exposure_fr, exposure_other), names_to = "language", values_to = "exposure") %>%
  group_by(language) %>%
  summarize(number = length(unique(subject_id)),
            mean_exposure = round(weightedMean(exposure, count), 1),
            sd_exposure = round(weightedSd(exposure, count), 1))


# Number of equal exposure to English and French
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant
  summarize(mean_exposure_eng = mean(exposure_eng),
            mean_exposure_fr = mean(exposure_fr),
            mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  # English exposure = French exposure
  filter(mean_exposure_eng == mean_exposure_fr) %>%
  # count how many 
  nrow()
```

# Maternal education level in years
```{r cache=TRUE}
clean_demog %>% 
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, primary_caregiver_eng, primary_caregiver_education_eng) %>% # keeping "primary_caregiver_education_eng" because all listed mother as the primary caregiver
  distinct(subject_id, .keep_all = TRUE) %>%
  group_by(primary_caregiver_eng) %>%
  summarize(mean_edu_level = mean(primary_caregiver_education_eng, na.rm=T),
            sd_edu_level = sd(primary_caregiver_education_eng, na.rm=T),
            min_edu_level = min(primary_caregiver_education_eng, na.rm=T),
            max_edu_level = max(primary_caregiver_education_eng, na.rm=T))
  
# how many mothers completed a university degree or higher (years 16 = university graduate, > 16 = advanced degree)
clean_demog %>% 
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, primary_caregiver_eng, primary_caregiver_education_eng) %>% # keeping "primary_caregiver_education_eng" because all listed mother as the primary caregiver
  distinct(subject_id, .keep_all = TRUE) %>%
  # university graduate? (1 = yes, 0 = no)
  mutate(university_degree = if_else(primary_caregiver_education_eng >= 16, 1, 0)) %>%
  # summary
  dplyr::count(university_degree) %>%
  summarize(university_degree = university_degree,
            n = n, 
            percentage = n/sum(n)*100)

```


# Who filled out the forms
```{r cache=TRUE}
# English CDIs
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, form_filler_eng) %>%
  filter(!is.na(form_filler_eng)) %>%
  dplyr::count(form_filler_eng) %>%
  summarize(form_filler_eng = form_filler_eng,
            n = n,
            percentage = n/sum(n)*100)

# French CDIs
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, form_filler_fr) %>%
  filter(!is.na(form_filler_fr)) %>%
  dplyr::count(form_filler_fr) %>%
  summarize(form_filler_fr = form_filler_fr,
            n = n,
            percentage = n/sum(n)*100)

# How often the forms in each language at each administration were filled out by the same parent?
## English CDIs
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, form_filler_eng) %>%
  #pivot_longer(form_filler_eng:form_filler_fr, names_to = "form_language", values_to = "filler") %>%
  pivot_wider(names_from = n_months, values_from = form_filler_eng) %>%
  # check row-wise and see if all values are identical
  rowwise %>%
  mutate(n_formfiller = as.factor(n_distinct(unlist(select(cur_data(), 2:11)), na.rm = TRUE))) %>% # if all consistent, then it'll return 1; otherwise the value would be > 1 indicating more than one form filler identity
  ungroup() %>%
  # summary
  dplyr::count(n_formfiller) %>%
  summarize(n_formfiller = n_formfiller,
            n = n,
            percentage = n/sum(n)*100)

# French CDIs
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, form_filler_fr) %>%
  #pivot_longer(form_filler_eng:form_filler_fr, names_to = "form_language", values_to = "filler") %>%
  pivot_wider(names_from = n_months, values_from = form_filler_fr) %>%
  # check row-wise and see if all values are identical
  rowwise %>%
  mutate(n_formfiller = as.factor(n_distinct(unlist(select(cur_data(), 2:11)), na.rm = TRUE))) %>% # if all consistent, then it'll return 1; otherwise the value would be > 1 indicating more than one form filler identity
  ungroup() %>%
  # summary
  dplyr::count(n_formfiller) %>%
  summarize(n_formfiller = n_formfiller,
            n = n,
            percentage = n/sum(n)*100)

```

# Primary caregiver in each language
```{r cache=TRUE}
# English CDIs
## percentage of whom is the primary caregiver in English
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  filter(!is.na(completed_eng)) %>%
  select(subject_id, primary_caregiver_eng) %>%
  dplyr::count(primary_caregiver_eng) %>%
  summarize(primary_caregiver_eng = primary_caregiver_eng,
            n = n,
            percentage = n/sum(n)*100)

# French CDIs
## percentage of whom is the primary caregiver in French
clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  filter(!is.na(completed_fr)) %>%
  select(subject_id, primary_caregiver_fr) %>%
  dplyr::count(primary_caregiver_fr) %>%
  summarize(primary_caregiver_fr = primary_caregiver_fr,
            n = n,
            percentage = n/sum(n)*100)
```

# Number of completed CDIs in each language
```{r cache=TRUE}
# Total number of completed CDIs
## Completed English CDIs
clean_demog  %>%
  filter(monolingual != 1) %>%
  select(c(subject_id, n_months, completed_eng)) %>%
  group_by(subject_id) %>%
  filter(!is.na(completed_eng)) %>%
  nrow()
  
## Completed French CDIs
clean_demog  %>%
  filter(monolingual != 1) %>%
  select(c(subject_id, n_months, completed_fr)) %>%
  group_by(subject_id) %>%
  filter(!is.na(completed_fr)) %>%
  nrow()

## Completed both English and French CDIs
clean_demog %>%
  filter(monolingual != 1) %>%
  select(c(subject_id, n_months, completed_both)) %>%
  group_by(subject_id) %>%
  filter(completed_both == TRUE) %>%
  nrow()

  
# Average number of completed CDIs
## English entries
clean_demog  %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  group_by(subject_id) %>%
  dplyr::count(completed_eng) %>%
  filter(!is.na(completed_eng)) %>%
  ungroup()%>%
  summarize(mean_entry_eng = mean(n, na.rm=T),
            sd_entry_eng = sd(n, na.rm=T),
            min_entry_eng = min(n, na.rm=T),
            max_entry_eng = max(n, na.rm=T))
            
## French entries
clean_demog  %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  group_by(subject_id) %>%
  dplyr::count(completed_fr) %>%
  filter(!is.na(completed_fr)) %>%
  ungroup()%>%
  summarize(mean_entry_fr = mean(n, na.rm=T),
            sd_entry_fr = sd(n, na.rm=T),
            min_entry_fr = min(n, na.rm=T),
            max_entry_fr = max(n, na.rm=T))
```


# Exclusion criteria
```{r cache=TRUE}
# Monolingual?
clean_demog %>%
  select(subject_id, monolingual) %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  filter(monolingual == 1) # returns 6

# Preterm?
clean_demog %>%
  select(subject_id, preterm) %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  filter(preterm == 1) # returns 0

# Underweight?
clean_demog %>%
  select(subject_id, underweight) %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  filter(underweight == 1) # returns 2

# Ear infection?
clean_demog %>%
  select(subject_id, ear_infections_problem, ear_infections_reason) %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  filter(ear_infections_problem == 1) # returns 0

# Hearing loss?
clean_demog %>%
  select(subject_id, hearing_loss_problem, hearing_loss_reason) %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  filter(hearing_loss_problem == 1) # returns 1; we could still include this data for now

# Vision problem?
clean_demog %>%
  select(subject_id, vision_problem, vision_reason) %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  filter(vision_problem == 1) # returns 2; we could still include the data for now

# Illness?
clean_demog %>%
  select(subject_id, illnesses_problem, illnesses_reason) %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  filter(illnesses_problem == 1) # returns 3; OK to include the data

# Service?
clean_demog %>%
  select(subject_id, services_problem, services_reason) %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  filter(services_problem == 1) # returns 0

# Worried?
clean_demog %>%
  select(subject_id, worried_problem, worried_reason) %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  filter(worried_problem == 1) # returns 3

# Learning disability
clean_demog %>%
  select(subject_id, learning_disability_problem, learning_disability_reason) %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  filter(learning_disability_problem == 1) # returns 12; OK to include the data

```


# Create a list of participants to be excluded
## (for now & Lori's thesis, we only care about language group: exclude monolinguals. We are ignoring the other criteria in this script, because--after checking--no infants should be excluded based on health criteria)
```{r cache=TRUE}
exclusions <- clean_demog %>%
  filter(monolingual == 1 | # exclude mono
           preterm == 1 | # exclude preterm
           underweight == 1) %>%  # exclude underweight
  distinct(subject_id)

write.csv(exclusions, here("data_keepers/exclusions.csv"), row.names=F)

```
