---
title: "Data dictionary: Cognates are advantaged in early bilingual expressive vocabulary development"
author: "Lori,Mitchell, Rachel K.Y. Tsui, & Krista Byers-Heinlein"
date: '`r format(Sys.time(), "%a %b %d %X %Y")`'
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: 'hide'
    self_contained: true
  pdf_document:
    toc: yes
    toc_depth: 4
    latex_engine: xelatex
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  warning = TRUE, # show warnings during codebook generation
  message = TRUE, # show messages during codebook generation
  error = TRUE, # do not interrupt codebook generation in case of errors,
                # usually better for debugging
  echo = TRUE  # show R code
)
ggplot2::theme_set(ggplot2::theme_bw())
pander::panderOptions("table.split.table", Inf)
```

```{r library, message=FALSE, warning=FALSE}
# load libraries
library(codebook)
library(here)
library(dplyr)
library(tidyverse)
library(future)
library(labelled)
```

This is a data dictionary for the data set used in the paper "Cognates are advantaged in early bilingual expressive vocabulary development".

```{r message=FALSE, warning=FALSE}
#load dataset
keepers_cognate_full <- rio::import(here::here("data_keepers/public_keepers_cognate_full.csv"))

keepers_cognate_matched <- rio::import(here::here("data_keepers/public_keepers_cognate_matched.csv"))
```

# Variables {.tabset}

## keepers_cognate_full: Overview
```{r}
codebook_items(keepers_cognate_full)
```

## keepers_cognate_matched: Overview
```{r}
codebook_items(keepers_cognate_matched)
```

## subject_id & n_months
* __subject_id__: Unique participant ID
* __n_months__: The month of administration, where 1 means the participant contributed data the first month after they entered the study. The subsequent numbers refers to the months that they contributed data along the study.

The dataset included N = `r length(unique(keepers_cognate_full$subject_id))` participants. 

```{r message=FALSE, warning=FALSE}
keepers_cognate_full %>% 
  mutate(administration = str_c(subject_id, "_", n_months)) %>%
  summarize(N_subject_id = n_distinct(subject_id),
            N_administration = n_distinct(administration))
```
## completed
* __completed_eng__: whether the English form was completed that month
* __completed_fr__: whether the French form was completed that month
* __completed_both__: whether both the English and French forms was completed that month

In the final analysis, we only kept data where both the English and French forms were completed within the same month.

```{r message=FALSE, warning=FALSE}
keepers_cognate_full %>% 
  mutate(administration = str_c(subject_id, "_", n_months)) %>%
  distinct(administration, .keep_all = TRUE) %>%
  summarize(n_completed_eng = length(completed_eng),
            n_completed_fr = length(completed_fr),
            n_completed_both = length(completed_both))
```

## age
* __age_days__: Age in days, calculated by _Date of test_ - _Date of birth_ (which are both removed from the public data to protect the participants' information)
* __age_months__: Age in months (with decimal), calculated by age_days/(365.25/12)

```{r message=FALSE, warning=FALSE}
keepers_cognate_full %>%
  summarize(mean_age_days = mean(age_days, na.rm = T),
            sd_age_days = sd(age_days, na.rm = T),
            min_age_days = min(age_days, na.rm = T),
            max_age_days = max(age_days, na.rm = T))
```

```{r message=FALSE, warning=FALSE}
keepers_cognate_full %>% 
  ggplot(aes(age_days)) +
  geom_histogram()
```

## sex
* __sex__: categorical with 2 levels: Female & Male

```{r message=FALSE, warning=FALSE}
keepers_cognate_full %>%
  distinct(subject_id, .keep_all = TRUE) %>% 
  count(sex) %>%
  mutate(percentage = round(n/sum(n)*100, 2)) 
```

## total_vocab
* __total_eng_vocab__: The total number of words the participant produced on the English CDI form within the complete cognate list (i.e., 537 translation equivalents)
* __total_fr_vocab__: The total number of words the participant produced on the French CDI form within the complete cognate list (i.e., 537 translation equivalents)
* __total_vocabulary__: The total number of vocabulary produced on both the English and French forms within the complete cognate list (i.e., 537 translation equivalents), calculated by _total_eng_vocab + total_fr_vocab_

```{r message=FALSE, warning=FALSE}
keepers_cognate_full %>%
  mutate(administration = str_c(subject_id, "_", n_months)) %>%
  distinct(administration, .keep_all = TRUE) %>%
  pivot_longer(c(matches("total_")), names_to = "type", values_to = "n_words") %>%
  group_by(type) %>%
  summarize(mean_n_words = mean(n_words, na.rm=T),
            sd_n_words = sd(n_words, na.rm=T),
            min_n_words = min(n_words, na.rm=T),
            max_n_words = max(n_words, na.rm=T))
```

## word_pairs
* __word_pairs__: The translation equivalent pair with the English word listed before the hyphen and the French word after
* __cognate_status__: Whether the translation equivalent pair is "cognate" or "non-cognate" (2 levels)
* __english_item_id & english_item__: The item in the English CDI form and its item number on the CDI form
* __french_item_id & french_item__: The item in the French CDI form and its item number on the CDI form
* __eng_acquired__: Whether the participant produced the English item, with 2 levels: 1 = produced and 0 = not yet produced
* __fr_acquired__: Whether the participant produced the French item, with 2 levels: 1 = produced and 0 = not yet produced
* __Eng_AoA__ (only in keepers_cognate_matched.csv): Age of acquisition for each English CDI word retrieved from Wordbankr
* __Fr_AoA__ (only in keepers_cognate_matched.csv): Age of acquisition for each French CDI word retrieved from Wordbankr

### Full list
#### Number of word pairs in the full list
```{r message=FALSE, warning=FALSE}
keepers_cognate_full %>%
  filter(subject_id == 53279 & n_months == 1) %>%
  summarize(n = length(word_pairs))
```

#### Number of cognates vs. non-cognates in the full list
```{r message=FALSE, warning=FALSE}
keepers_cognate_full %>%
  filter(subject_id == 53279 & n_months == 1) %>%
  group_by(cognate_status) %>%
  summarize(n = length(cognate_status))
```

#### Number of English and French items in the full list
```{r message=FALSE, warning=FALSE}
keepers_cognate_full %>%
  filter(subject_id == 53279 & n_months == 1) %>%
  summarize(eng_n = length(english_item),
            fr_n = length(french_item))
```

### Matched list
#### Number of word pairs in the matched list
```{r message=FALSE, warning=FALSE}
keepers_cognate_matched %>%
  filter(subject_id == 53279 & n_months == 1) %>%
  summarize(n = length(word_pairs))
```

#### Number of cognates vs. non-cognates in the matched list
```{r message=FALSE, warning=FALSE}
keepers_cognate_matched %>%
  filter(subject_id == 53279 & n_months == 1) %>%
  group_by(cognate_status) %>%
  summarize(n = length(cognate_status))
```

#### Number of English and French items in the matched list
```{r message=FALSE, warning=FALSE}
keepers_cognate_matched %>%
  filter(subject_id == 53279 & n_months == 1) %>%
  summarize(eng_n = length(english_item),
            fr_n = length(french_item))
```

