---
title: "Cognates are advantaged in early bilingual expressive vocabulary development"
shorttitle: "Cognate advantage in bilingual infants"

header-includes:
  - \usepackage{amsmath}
  - \usepackage[labelformat=empty]{caption}
  - \usepackage{caption}
  - \usepackage[extra]{tipa}
  - \renewcommand{\topfraction}{1}
  - \renewcommand{\bottomfraction}{1}
  - \renewcommand{\textfraction}{.1}
  - \renewcommand{\floatpagefraction}{1}
  - \setcounter{topnumber}{9}
  - \setcounter{bottomnumber}{9}
  - \setcounter{totalnumber}{20}
  - \setcounter{dbltopnumber}{9}

author: 
  - name          : "Lori Mitchell"
    affiliation   : "1"
    email         : "lorimitchell7@hotmail.com"
  - name          : "Rachel Ka-Ying Tsui"
    affiliation   : "1" 
    email         : "rachelkytsui@gmail.com"
    corresponding : yes
    address       : "Department of Psychology, 7141 Sherbrooke St. West, Montreal, QC, Canada, H2T1V2. Rachel Ka-Ying Tsui is now at Laboratory for Language Development, RIKEN Center for Brain Science, 2-1 Hirosawa, Wako-shi, Saitama, Japan, 351-0198"
  - name          : "Krista Byers-Heinlein"
    affiliation   : "1" 
    email         : "k.byers@concordia.ca"
    address       : "Department of Psychology, 7141 Sherbrooke St. West, Montreal, QC, Canada, H2T1V2."


affiliation:
  - id            : "1"
    institution   : "Concordia University"

abstract: |
  Bilingual infants grow up with the unique experience of needing to learn two words for most concepts. These words are called translation equivalents, and translation equivalents that also sound similar (e.g., banana-banane) are called cognates. Research has consistently shown that children and adults process and name cognates more easily than non-cognates. The present study explored if there is such an advantage for cognate production in bilinguals’ early vocabulary development. Using longitudinal expressive vocabulary data collected from 47 English–French bilingual infants and toddlers starting at the ages of 16–20 months up to 27 months (a total of 219 monthly administrations in both English and French), results showed that overall children produced a greater proportion of cognate words than non-cognate words on the MacArthur-Bates Communicative Development Inventories. The findings suggest that cognate learning is facilitated in early bilingual vocabulary development. Just as in monolingual infants, these results suggest that phonological overlap supports bilingual language acquisition.
 

keywords          : "bilingualism, infants, cognates, translation equivalents, phonological similarity, expressive vocabulary development"

wordcount         : 

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no # "yes" to omit identifying information from the title page

bibliography      : ["CogVocab_paper_references.bib"]
replace_ampersands: yes

class             : "man"
output            : papaja::apa6_docx # change to apa6_docx for word document
citation_package  : biblatex
nocite: '@*' # include all the references in the .bib file, because in-text citations were mostly typed manually to fit the apa7 style
---

\captionsetup[table]{labelformat=empty}

```{r setup, include=FALSE, cache=TRUE}
library(rmarkdown)
library(knitr)
library(papaja)
library(kableExtra)
library(english)
library(linguisticsdown)
library(here)
library(tidyverse)
library(tidylog)
library(wordbankr)
library(matrixStats)
library(lme4)
library(lmerTest)
library(ggplot2)
library(patchwork)
library(gridExtra)
library("ggpubr")
library(broom)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)

```

```{r load_data, cache = TRUE}
# read in the final cleaned keepers_data
clean_demog <- read.csv(here::here("data_keepers/public_clean_demog.csv"))

exclusion <- read.csv(here::here("data_keepers/exclusions.csv"))

public_keepers_cognate_full <- read.csv(here::here("data_keepers/public_keepers_cognate_full.csv"))

public_keepers_cognate_matched <- read.csv(here::here("data_keepers/public_keepers_cognate_matched.csv"))
```

Infants understand some words during the first year of life, and begin to produce words around their first birthday (Fenson et al., 2007). To do so, infants must represent both the phonological and semantic aspects of a word, and associate the two. Intriguingly, the network of words that children already know appears to shape the words that they will learn: monolingual infants are more likely to learn words that are phonologically (Luce & Pisoni, 1998; Coady & Aslin, 2003) and semantically (Coady & Aslin, 2003) similar to those they already know. Bilingual infants provide a unique perspective into understanding children’s developing lexical networks, as they must acquire translation equivalents, which are cross-language synonyms with complete or nearly complete semantic overlap (e.g., “apple” in English and _“pomme”_ in French; De Houwer et al., 2006; Legacy et al., 2017; Pearson et al., 1995; see White et al. (2017) for a discussion of convergence in bilinguals’ semantic representations). For infants acquiring typologically or historically related languages, some of these translation equivalents will be cognates, which are also phonologically similar (e.g., “banana” \textipa{/b@n\ae{}n@/} in English and _“banane”_ \textipa{/b\textscripta{}n\textscripta{}n/} in French). This current study aimed to understand the impact of cognate status on the acquisition of words in young bilinguals by examining whether bilingual infants produce cognates more readily than non-cognates in early language development.

## Translation Equivalents
Translation equivalents are an important part of early language development for bilingual children. While early researchers claimed that bilinguals avoid learning translation equivalents (Volterra & Taeschner, 1978), recent work shows that bilingual infants acquire translation equivalents from an early age (Legacy et al., 2017; Pearson et al., 1995). Bilingual infants begin to produce translation equivalents by 16 months, and produce more translation equivalents with age as their vocabulaires grow (Legacy et al., 2017). The strong semantic overlap of a word in one language seems to facilitate the acquisition of its translation equivalent in the other language, at least at younger ages when bilingual infants have smaller vocabularies (Bilson et al., 2015; Tsui et al., in press). By the age of 27 months, bilingual toddlers recognize a target word more accurately when preceded by its translation equivalent (Floccia et al., 2020). 

For bilingual infants, some translation equivalents sound very similar. Specifically, cognates are a type of translation equivalents that have significant phonological overlap, typically due to a shared etymology [^1]. Cognates range in their degree of phonological similarity: For example, English “banana” \textipa{/b@n\ae{}n@/} and French _“banane”_ \textipa{/b\textscripta{}n\textscripta{}n/} are highly phonetically similar, while English “pants” \textipa{/p\ae{}nts/} and French _“pantalon”_ \textipa{/p\~A tAl\~O/} are more different including a different number of syllables. Some typologically close languages even have form-identical cognates, such as the word _“sí”_ \textipa{/si/} which means “yes” in both Spanish and Catalan.

[^1]: Cognates can also overlap in their orthography, but we do not address orthography in this paper as our participants were too young to read.

Cognates appear to have a special status in bilingual language processing and production. Previous research has reported a cognate facilitation effect where bilinguals are better and quicker at identifying and naming cognates than non-cognates when performing vocabulary tasks (e.g., Costa et al., 2000; Kelley & Kohnert, 2012; Sheng et al., 2016). This type of advantage for cognates has been reported in bilingual adults (e.g., Costa et al., 2000) as well as in school-aged children (e.g., Kelley & Kohnert, 2012; Sheng et al., 2016). For example, Kelley and Kohnert (2012) provide evidence for the cognate facilitation effect in Spanish-speaking English learners between the ages of 8 and 13 years old, where children identified and named more cognates than non-cognates in receptive and expressive vocabulary tasks. A similar cognate advantage has been found for picture naming and translation tasks for English–Spanish and English–German 4- to 8-year-old children where bilingual children were more accurate in naming cognates and faster at translating cognates than non-cognates (Schelletter, 2002; Sheng et al., 2016). Therefore, cognates seem to be advantaged in school-aged bilingual children’s language processing and production.

## Effects of Phonological Similarity on Early Word Learning
The advantage for cognates could be attributed to the phonological overlap between words, which may make them easier to learn. Existing literature on monolinguals has reported that children are more likely to produce words that sound similar to other words in their lexicons (e.g., “at” and “cat”, “hat” and “cat”), especially at younger ages (e.g., Jones & Brandt, 2019). For instance, looking at 300 British English-speaking children aged 12 to 25 months, Jones and Brandt (2019) found that the strength of phonological similarity between words was an important predictor for word production (but not comprehension), whereby young children tended to produce words that follow similar phonological patterns. Similarly, using archival expressive vocabulary data from 1,800 16- to 30-month-old American infants, it was shown that infants produced more nouns with many phonological neighbours than those with few phonological neighbours (Storkel, 2009). It is possible that the high degree of phonological similarity aids word acquisition through sounds already established in the lexicons. For example, Demke et al. (2002) found that hearing real-word phonological neighbours facilitated the learning of new pseudowords. Another possibility is that the words that share a high degree of phonological similarity in the language input are learned first by infants, as supported by a recent study looking at the developing lexicons of young infants across 10 languages (Fourtassi et al., 2020). Overall, learning a new word with close phonological neighbours seems to help learners maintain the new word in memory, making similar-sounding words easier to acquire and produce (e.g., Coady & Aslin, 2003; Demke et al., 2002; Jones & Brandt, 2019).

Extending this notion to bilingual infants, some evidence suggests that phonological similarity facilitates vocabulary learning across languages as well. For example, Gampe et al. (2021) examined parent-reported vocabulary size of 18- to 36-month-old children learning Swiss German and another language. Children learning languages that were more phonologically similar to Swiss German (e.g., standard German, Dutch, English) produced more words than children learning languages that were more phonologically dissimilar (e.g., Turkish, French). Moreover, children learning more similar languages learned more cognate translation equivalents, while the number of non-cognate translation equivalents was similar across groups. These results are consistent with other studies reporting that language distance affects early bilingual language acquisition (e.g., Blom et al., 2019; Gampe et al., 2021; Havy et al., 2016; Sheng et al., 2016).

However, not all studies have reported a generalized advantage for cognates in vocabulary learning. In a study of younger children, Bosch and Ramon-Casas (2014) used parent reports to examine word production in 18-month-olds learning Spanish and Catalan, two strongly related languages that share many form-identical (e.g, “yes” is _“sí”_ \textipa{/si/} in both Spanish and Catalan) and form-similar (e.g., “hand” is _“mano”_ \textipa{/mAno/} in Spanish and _“mà”_ \textipa{/mA/} in Catalan) cognates. Results indicated that 28% of the words produced by the bilingual infants were form-identical cognates, while less than 2% of words were form-similar cognates or non-cognate translation equivalents (Bosch & Ramon-Casas, 2014). One explanation for this finding is that for form-identical cognates, infants only need to learn a single form for a particular concept, which they can then transfer across their languages. Based on these results, bilingual infants may not benefit from cognates’ phonological overlap unless that overlap is perfect. Indeed, there is some evidence that Spanish–Catalan infants are somewhat insensitive to phonological distinctions in form-similar cognates (Ramon-Casas et al., 2009; Ramon-Casas & Bosch, 2010), perhaps even representing them as form-identical. Another interpretation of this result is that the effect of cognates on bilingual vocabulary learning changes across development, which could explain the discrepant results of the 18-month-old sample studied by Bosch & Ramon-Casas (2014), and the 18- to 36-month-old sample studied by Gampe et al. (2021).

## Current study
To better understand the impact of phonological overlap on bilingual infants’ vocabulary learning, we examined the production of cognate and non-cognate translation equivalents in French–English bilingual infants. English and French share many form-similar cognates due to historical language contact (Choi, 2019), although only a few form-identical cognates. Despite the presence of cognates, note that these two languages belong to different language families: English is a Germanic language and French is a Romance language. Previous work looked at learners of closely related languages with many form-identical cognates (Spanish and Catalan; Ramon-Casas & Bosch, 2010), or else a heterogeneous group of bilinguals learning many different language pairs (Gampe et al., 2021). Thus, our study provided an important test of the generalizability of these results in a new and homogeneous population of young bilinguals.

We collected monthly vocabulary data on French–English bilingual infants’ word production starting when children were between the ages of 16–20 months and ending when were up to 27 months of age using the MacArthur-Bates Web-Communicative Developmental Inventory: Words and Sentences form in American English (Fenson et al., 2007) and Québec French (Trudeau et al., 1999). Uniquely, our dataset was longitudinal, allowing us to investigate potential developmental effects. We focused our analysis on translation equivalent pairs and then classified the pairs according to cognate status (cognate or non-cognate words). We counted children’s production of both translation equivalent pairs (e.g., whether they produced both “apple” \textipa{/\ae{}p@l/} and _“pomme”_ \textipa{/p\textopeno{}m/}, or both “banana” \textipa{/b@n\ae{}n@/} and _“banane”_ \textipa{/b\textscripta{}n\textscripta{}n/}), as well as individual words independent of whether children produced its translation equivalent. Since it is not possible to randomly assign our main variable of interest (cognate status), we analyzed both a complete list of cognate and non-cognate words, as well as a carefully selected subset of these cognate and non-cognate words which were matched on age of acquisition and on word category (e.g., words about food) where possible. 

We hypothesized that French–English bilinguals would more readily produce cognates than non-cognates. Thus, we predicted that French–English bilingual infants would produce proportionally more translation equivalent words and pairs that were cognates than non-cognates. We likewise anticipated an interaction between cognate status and age, with a stronger effect of cognate status at older ages as the infants’ vocabulary size (and the number of translation equivalent words and pairs produced) grew.

# Method
The present research was approved by the Human Research Ethics Committee at Concordia University [certification #10000439]. Participation was on a voluntary basis and the families were free to withdraw at any time. The study design was pre-registered at https://osf.io/rh7av.

## Participants
```{r include=FALSE, cache=TRUE}
# Number of total participants (after removing incompleted entries, monolinguals, and pilot data)
total_n <- clean_demog %>%
  filter(monolingual != 1) %>%
  distinct(subject_id) %>%
  nrow()


# Number of girls 
n_female <- clean_demog %>%
  filter(monolingual != 1) %>%
  select(subject_id, sex) %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  filter(sex == "Female") %>%
  nrow()


# Age range for 1st month data entry
mean_age_1st_entry <- clean_demog %>%
  filter(monolingual != 1) %>%
  filter(n_months == 1) %>%
  summarize(mean_age = mean(age_months, na.rm=T),
            sd_age = sd(age_months, na.rm=T),
            min_age = min(age_months, na.rm=T),
            max_age = max(age_months, na.rm=T))


# Age range for final month data entry
mean_age_final_entry <- clean_demog %>%
  filter(monolingual != 1) %>%
  arrange(subject_id, n_months) %>%
  # get the last row of data within each participant
  group_by(subject_id) %>%
  slice(n()) %>%
  ungroup() %>%
  # summarize
  summarize(mean_age = mean(age_months, na.rm=T),
            sd_age = sd(age_months, na.rm=T),
            min_age = min(age_months, na.rm=T),
            max_age = max(age_months, na.rm=T)) # the oldest age is 27m


# Total number of completed CDIs
## Completed English CDIs
n_eng_administration <- clean_demog  %>%
  filter(monolingual != 1) %>%
  select(c(subject_id, n_months, completed_eng)) %>%
  group_by(subject_id) %>%
  filter(!is.na(completed_eng)) %>%
  nrow()
  
## Completed French CDIs
n_fr_administration <- clean_demog  %>%
  filter(monolingual != 1) %>%
  select(c(subject_id, n_months, completed_fr)) %>%
  group_by(subject_id) %>%
  filter(!is.na(completed_fr)) %>%
  nrow()

## Completed both English and French CDIs
n_both_administration <- clean_demog %>%
  filter(monolingual != 1) %>%
  select(c(subject_id, n_months, completed_both)) %>%
  group_by(subject_id) %>%
  filter(completed_both == TRUE) %>%
  nrow()


# Number of participants who contributed data more than one time point
n_multiple_entry <- clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  group_by(subject_id) %>%
  dplyr::count(subject_id) %>%
  filter(n > 1) %>%
  nrow()


# Average number of completed CDIs
## English entries
n_avg_entries <- clean_demog  %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  group_by(subject_id) %>%
  dplyr::count(completed_eng) %>%
  filter(!is.na(completed_eng)) %>%
  ungroup()%>%
  summarize(mean_entry = mean(n, na.rm=T),
            sd_entry = sd(n, na.rm=T),
            min_entry = min(n, na.rm=T),
            max_entry = max(n, na.rm=T))
```

The current study comprised data from `r total_n` French–English bilingual infants (`r n_female` females) collected from August 2020 to May 2021, as part of a larger ongoing longitudinal study which aims to collect data from 100 bilingual infants. Participating infants were aged between 16 and 20 months at the onset of participation (mean starting age = `r round(mean_age_1st_entry$mean_age,2)` months, SD = `r round(mean_age_1st_entry$sd_age,2)`, range = `r round(mean_age_1st_entry$min_age,2)` – `r round(mean_age_1st_entry$max_age,2)`), and were aged between `r round(mean_age_final_entry$min_age,0)` to `r round(mean_age_final_entry$max_age,0)` months at their final time of participation (M = `r round(mean_age_final_entry$mean_age,2)` months, SD = `r round(mean_age_final_entry$sd_age,2)`, range = `r round(mean_age_final_entry$min_age,2)` – `r round(mean_age_final_entry$max_age,2)`). Participants were recruited from Québec, Canada through government birth lists, social media, and participating families’ referrals. Inclusion criteria were the following: full-term pregnancy (i.e., at least 37 weeks of gestation), normal birth weight (> 2500 grams), and no reported developmental delays or any hearing or vision problems. Bilingual infants were defined as those exposed to each of English and French for at least 10% and at most 90% of the time over the course of their lives since birth, with less than 10% of exposure to a third language. To capture a wider range of bilingual experience, the language exposure range in this study was broader than some studies (e.g., Morin-Lessard & Byers-Heinlein, 2019; Sebastián-Gallés & Bosch, 2009) but similar to the range used in others (e.g., Hoff & Ribot, 2017; Place & Hoff, 2011). 

```{r include=FALSE, cache=TRUE}
# Average language exposure (weighted mean is calculated as participants contributed different number of LEQ update data)
## weighted mean across all participants
mean_langexp <- clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  mutate(count = n()) %>%
  ungroup() %>%
  pivot_longer(c(exposure_eng, exposure_fr, exposure_other), names_to = "language", values_to = "exposure") %>%
  group_by(language) %>%
  summarize(mean_exposure = round(weightedMean(exposure, count), 1),
            sd_exposure = round(weightedSd(exposure, count), 1))

## minimum and maximum of individual averages
range_langexp <- clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant
  summarize(mean_exposure_eng = mean(exposure_eng),
            mean_exposure_fr = mean(exposure_fr),
            mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  # minimum and maximum of the individual averages
  pivot_longer(c(mean_exposure_eng, mean_exposure_fr, mean_exposure_other), names_to = "language", values_to = "mean_exposure") %>%
  group_by(language) %>%
  summarize(min_exposure = min(mean_exposure),
            max_exposure = max(mean_exposure))


# Number of English dominant
## weighted mean across all participants
mean_engdom <- clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant to define dominance
  mutate(count = n(),
         mean_exposure_eng = mean(exposure_eng),
         mean_exposure_fr = mean(exposure_fr),
         mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  filter(mean_exposure_eng > mean_exposure_fr) %>% # English exposure > French exposure
  select(-c(matches("mean"))) %>%
  # calculating weigthed mean
  pivot_longer(c(exposure_eng, exposure_fr, exposure_other), names_to = "language", values_to = "exposure") %>%
  group_by(language) %>%
  summarize(number = length(unique(subject_id)),
            mean_exposure = round(weightedMean(exposure, count), 1),
            sd_exposure = round(weightedSd(exposure, count), 1))

## minimum and maximum of individual averages
range_engdom <-clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant
  summarize(mean_exposure_eng = mean(exposure_eng),
            mean_exposure_fr = mean(exposure_fr),
            mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  # English exposure > French exposure
  filter(mean_exposure_eng > mean_exposure_fr) %>%
  # minimum and maximum of the individual averages
  pivot_longer(c(mean_exposure_eng, mean_exposure_fr, mean_exposure_other), names_to = "language", values_to = "mean_exposure") %>%
  group_by(language) %>%
  summarize(number = length(unique(subject_id)),
            min_exposure = min(mean_exposure),
            max_exposure = max(mean_exposure))


# Number of French dominant
## weighted mean across all participants
mean_frdom <- clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant to define dominance
  mutate(count = n(),
         mean_exposure_eng = mean(exposure_eng),
         mean_exposure_fr = mean(exposure_fr),
         mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  filter(mean_exposure_eng < mean_exposure_fr) %>% # English exposure < French exposure
  select(-c(matches("mean"))) %>%
  # calculating weigthed mean
  pivot_longer(c(exposure_eng, exposure_fr, exposure_other), names_to = "language", values_to = "exposure") %>%
  group_by(language) %>%
  summarize(number = length(unique(subject_id)),
            mean_exposure = round(weightedMean(exposure, count), 1),
            sd_exposure = round(weightedSd(exposure, count), 1))

## minimum and maximum of individual averages
range_frdom <- clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant
  summarize(mean_exposure_eng = mean(exposure_eng),
            mean_exposure_fr = mean(exposure_fr),
            mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  # English exposure < French exposure
  filter(mean_exposure_eng < mean_exposure_fr) %>%
  # minimum and maximum of the individual averages
  pivot_longer(c(mean_exposure_eng, mean_exposure_fr, mean_exposure_other), names_to = "language", values_to = "mean_exposure") %>%
  group_by(language) %>%
  summarize(number = length(unique(subject_id)),
            min_exposure = min(mean_exposure),
            max_exposure = max(mean_exposure))


# Number of equal exposure to English and French
n_equalexp <- clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant
  summarize(mean_exposure_eng = mean(exposure_eng),
            mean_exposure_fr = mean(exposure_fr),
            mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  # English exposure = French exposure
  filter(mean_exposure_eng == mean_exposure_fr) %>%
  # count how many 
  nrow()


# Maternal education level in years
mean_maternaledu <- clean_demog %>% 
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, primary_caregiver_eng, primary_caregiver_education_eng) %>% # keeping "primary_caregiver_education_eng" because all listed mother as the primary caregiver
  distinct(subject_id, .keep_all = TRUE) %>%
  group_by(primary_caregiver_eng) %>%
  summarize(mean_edu_level = mean(primary_caregiver_education_eng, na.rm=T),
            sd_edu_level = sd(primary_caregiver_education_eng, na.rm=T),
            min_edu_level = min(primary_caregiver_education_eng, na.rm=T),
            max_edu_level = max(primary_caregiver_education_eng, na.rm=T))
  
# how many mothers completed a university degree or higher (years 16 = university graduate, > 16 = advanced degree)
n_maternal_uni <- clean_demog %>% 
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, primary_caregiver_eng, primary_caregiver_education_eng) %>% # keeping "primary_caregiver_education_eng" because all listed mother as the primary caregiver
  distinct(subject_id, .keep_all = TRUE) %>%
  # university graduate? (1 = yes, 0 = no)
  mutate(university_degree = if_else(primary_caregiver_education_eng >= 16, 1, 0)) %>%
  # summary
  dplyr::count(university_degree) %>%
  summarize(university_degree = university_degree,
            n = n, 
            percentage = n/sum(n)*100)
```

In total, parents completed `r n_eng_administration` English CDI administrations and `r n_fr_administration` French CDI administrations. We retained only cases where both the English and French were completed at the same time point to be able to determine infants’ translation equivalent knowledge. This left us with `r n_both_administration` completed administrations from `r length(unique(public_keepers_cognate_full$subject_id))` infants. `r english::Words(length(unique(public_keepers_cognate_full$subject_id)) - n_multiple_entry)` infants contributed data at only one time point, and `r n_multiple_entry` infants contributed data at more than one time point, with participants contributing an average of `r round(n_avg_entries$mean_entry,1)` measurements for each language (SD = `r round(n_avg_entries$sd_entry,2)`, range = `r n_avg_entries$min_entry` – `r n_avg_entries$max_entry`). On average across the `r n_both_administration` administrations, participating infants were exposed to English `r round(mean_langexp[1,2],1)`% of the time (SD = `r round(mean_langexp[1,3],1)`, range = `r round(range_langexp[1,2],0)` – `r round(range_langexp[1,3],0)`), to French `r round(mean_langexp[2,2],1)`% of the time (SD = `r round(mean_langexp[2,3],1)`, range = `r round(range_langexp[2,2],0)` – `r round(range_langexp[2,3],0)`), and to a third language `r round(mean_langexp[3,2],1)`% of the time (SD = `r round(mean_langexp[3,3],1)`, range = `r round(range_langexp[3,2],0)` – `r round(range_langexp[3,3],0)`). Of the `r length(unique(public_keepers_cognate_full$subject_id))` bilingual infants, `r mean_engdom[1,2]` were English-dominant (M = `r round(mean_engdom[1,3],1)`% English exposure, SD = `r round(mean_engdom[1,4],1)`, range = `r round(range_engdom[1,3],0)` – `r round(range_engdom[1,4],0)`), `r mean_frdom[1,2]` were French-dominant (M = `r round(mean_frdom[2,3],1)`% French exposure, SD = `r round(mean_frdom[2,4],1)`, range = `r round(range_frdom[2,3],0)` – `r round(range_frdom[2,4],0)`), and `r n_equalexp` reported equal exposure to both English and French. The average maternal education level was `r round(mean_maternaledu[1,2],2)` years (SD = `r round(mean_maternaledu[1,3],2)`, range = `r round(mean_maternaledu[1,4],2)` – `r round(mean_maternaledu[1,5],2)`), and `r round(n_maternal_uni[2,3],1)`% of the mothers had completed a university degree or higher.

## Measures
### Web-based MacArthur-Bates Communicative Development Inventory: Words and Sentences (Web-CDI)
```{r include=FALSE, cache=TRUE}
# Who filled out the forms
## English CDIs
filledby_eng <- clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, form_filler_eng) %>%
  filter(!is.na(form_filler_eng)) %>%
  dplyr::count(form_filler_eng) %>%
  summarize(form_filler_eng = form_filler_eng,
            n = n,
            percentage = n/sum(n)*100)

## French CDIs
filledby_fr <- clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, form_filler_fr) %>%
  filter(!is.na(form_filler_fr)) %>%
  dplyr::count(form_filler_fr) %>%
  summarize(form_filler_fr = form_filler_fr,
            n = n,
            percentage = n/sum(n)*100)


# How often the forms in each language at each administration were filled out by the same parent?
## English CDIs
formfiller_eng <- clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, form_filler_eng) %>%
  #pivot_longer(form_filler_eng:form_filler_fr, names_to = "form_language", values_to = "filler") %>%
  pivot_wider(names_from = n_months, values_from = form_filler_eng) %>%
  # check row-wise and see if all values are identical
  rowwise %>%
  mutate(n_formfiller = as.factor(n_distinct(unlist(select(cur_data(), 2:11)), na.rm = TRUE))) %>% # if all consistent, then it'll return 1; otherwise the value would be > 1 indicating more than one form filler identity
  ungroup() %>%
  # summary
  dplyr::count(n_formfiller) %>%
  summarize(n_formfiller = n_formfiller,
            n = n,
            percentage = n/sum(n)*100)

# French CDIs
formfiller_fr <- clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, form_filler_fr) %>%
  #pivot_longer(form_filler_eng:form_filler_fr, names_to = "form_language", values_to = "filler") %>%
  pivot_wider(names_from = n_months, values_from = form_filler_fr) %>%
  # check row-wise and see if all values are identical
  rowwise %>%
  mutate(n_formfiller = as.factor(n_distinct(unlist(select(cur_data(), 2:11)), na.rm = TRUE))) %>% # if all consistent, then it'll return 1; otherwise the value would be > 1 indicating more than one form filler identity
  ungroup() %>%
  # summary
  dplyr::count(n_formfiller) %>%
  summarize(n_formfiller = n_formfiller,
            n = n,
            percentage = n/sum(n)*100)
```

The number of words produced in English and French was obtained monthly via the web-based versions of the MacArthur-Bates Web-Communicative Development Inventories: Words and Sentences form (Web-CDI; https://webcdi.stanford.edu/), using the American English version (Fenson et al., 2007) and the Québec French adaptation (“Mots et Énoncés”; Trudeau et al., 1999). Our study focused on the vocabulary checklist component of the CDIs, with 680 words in the English version and 664 words in the Québec French version. We asked the caregiver most familiar with the infant’s vocabulary in each language to complete the respective version, although following the instructions on the Web-CDI they could seek help from others who often speak the corresponding language with the infant. The English forms were completed by mothers (`r round(filledby_eng[which(filledby_eng$form_filler_eng=='Mother'),3],0)`%), fathers (`r round(filledby_eng[which(filledby_eng$form_filler_eng=='Father'),3],0)`%), and both parents (`r round(filledby_eng[which(filledby_eng$form_filler_eng=='Both caregivers'),3],0)`%), whereas the French forms were completed by mothers (`r round(filledby_fr[which(filledby_fr$form_filler_fr=='Mother'),3],0)`%), fathers (`r round(filledby_fr[which(filledby_fr$form_filler_fr=='Father'),3],0)`%), and both parents (`r round(filledby_fr[which(filledby_fr$form_filler_fr=='Both caregivers'),3],0)`%). Thus, most of the time, the same caregiver (usually the mother) filled out both forms. Generally, whichever caregiver completed forms in a particular language did so throughout the study, with the exception of `r formfiller_eng[which(formfiller_eng$n_formfiller!=1),2]` participants (`r round(formfiller_eng[which(formfiller_eng$n_formfiller!=1),3],1)`%) whose English forms were filled out by different caregivers for some administrations, and `r formfiller_fr[which(formfiller_fr$n_formfiller!=1),2]` participants (`r round(formfiller_fr[which(formfiller_fr$n_formfiller!=1),3],1)`%) whose French forms were filled out by different caregivers for some administrations. Infants’ demographic information including age and sex was also collected at the start of the Web-CDI.

### Language Exposure Questionnaire (LEQ) using the Multilingual Approach to Parent Language Estimates (MAPLE) 
The infant’s language exposure and background was measured with an adaptation of the Language Exposure Questionnaire (LEQ; Bosch & Sebastián-Gallés, 2001), using the Multilingual Approach to Parent Language Estimates (MAPLE; Byers-Heinlein et al., 2020). During a 15- to 20-minute structured interview, the primary caregiver(s) were asked questions about the infant’s language exposure from birth until their current age. This provided a global estimate of the percentage of exposure that the infant had to each of their languages across all contexts. 

## Procedure
Data collection for this study began in August 2020 and ended in May 2021, although the start date of participation varied across participants. On the first day of each month, links to the English and French Web-CDI forms were sent to the caregivers by email. On the forms, the words that were checked off in previous months were automatically filled in the following months; thus, caregivers only needed to check off the new words that their child produced each month. This was intended to reduce the burden on participants, and increase the response rate. We asked that the Web-CDI forms be completed during the first week of each month. A reminder was sent on the 8th of the month, and an extra week was given for caregivers who had not yet completed the forms. If caregivers still did not complete the form, they were asked to resume their participation the following month. Once the forms were completed, caregivers received a brief report about their child’s vocabulary knowledge at that time point, including the total number of words that their child produced as well as the breakdown of the categories (such as animals, food, furniture, etc.) for which their child produced words. 

At the first data collection time point, caregivers also completed the LEQ questionnaire with a trained research assistant over the online video chat application Zoom. This was repeated every five months to track any potential changes in the infant’s language exposure. 

## Identification of Translation Equivalents and Cognates 
```{r include=FALSE, cache=TRUE}
CognateList_full_611TE <- read.csv(here::here("01_cognate_lists/CognateList_full_611TE.csv")) %>%
  dplyr::count(cognate_status)
```
A list of translation equivalents on the English and French forms of the CDI was created by three proficient English–French bilingual adults who carefully examined the English and French versions of the CDIs; a total of 611 translation equivalent pairs were identified (the full list is available at https://osf.io/7fz6c/; Gonzalez-Barrero et al., 2020). Next, bilingual research assistants identified `r CognateList_full_611TE[which(CognateList_full_611TE$cognate_status=="cognate"),2]` of the possible 611 translation equivalent pairs as cognates, with the remaining `r CognateList_full_611TE[which(CognateList_full_611TE$cognate_status=="non-cognate"),2]` words as non-cognates. Phonological similarity of the identified cognates were further confirmed by bilingual undergraduate students who were asked to evaluate the phonological overlap of recordings of those words. This method was preferred to other methods that focus on orthography (overlap in spelling), since infants acquire language through spoken words as opposed to reading. These steps were carried out in the Concordia Infant Research Laboratory for different projects, prior to the current study. 

```{r include=FALSE, cache=TRUE}
CognateList_full_clean <- read.csv(here::here("01_cognate_lists/CognateList_full_clean.csv")) 

n_CognateList_full_clean <- CognateList_full_clean %>%
  dplyr::count(cognate_status)
```
From the list of 611 translation equivalents, we further excluded any translation equivalent pairs that had complex relationships rather than one-to-one mappings. For example, “noodle” forms a translation equivalent pair with either the French word _“nouilles”_ or _“pâtes”_, where both French words are listed together as one item on the French CDI form. These pairs were removed because we could not know which form (e.g., _“nouilles”_ or _“pâtes”_) the infant produced, and we were not able to classify these pairs as either cognates or non-cognates. This left a complete list of `r nrow(CognateList_full_clean)` translation equivalents (`r nrow(CognateList_full_clean[which(CognateList_full_clean$cognate_status=="cognate"),])` cognates and `r nrow(CognateList_full_clean[which(CognateList_full_clean$cognate_status=="non-cognate"),])` non-cognates). 

Note that the cognates and non-cognates in this list could vary systematically on correlated factors including variations in parts of speech and differences in age of acquisition of certain words between languages. Within the full list of translation equivalents, we thus identified a matched subset of cognates and non-cognates.

```{r include=FALSE, cache=TRUE}
CognateList_nouns <- read.csv(here::here("01_cognate_lists/CognateList_nouns.csv"))

CognateList_nouns_AoA <- read.csv(here::here("01_cognate_lists/CognateList_nouns_AOA.csv"))

CognateList_matching <- read.csv(here::here("01_cognate_lists/CognateList_matching.csv")) %>%
  mutate(exact_match = case_when(startsWith(as.character(matches), "1.1") ~ "non-exact",
                                 TRUE ~ "exact"))
```

The matched list was first restricted to nouns as infants show a noun bias in language acquisition (Caselli et al., 1995), and doing so matched the cognates and non-cognates for part of speech. Next, the remaining `r nrow(CognateList_nouns)` translation equivalents (cognates = `r nrow(CognateList_nouns[which(CognateList_nouns$cognate_status=="cognate"),])`, non-cognates = `r nrow(CognateList_nouns[which(CognateList_nouns$cognate_status=="non-cognate"),])`) were matched on age of acquisition and word category where possible (e.g., food, furniture, etc.). However, data on age of acquisition, which was obtained from the wordbankr package (Version 0.3.1; Frank et al., 2017), was not available for `r nrow(CognateList_nouns) - nrow(CognateList_nouns_AoA)` translation equivalents which were therefore removed, leaving `r nrow(CognateList_nouns_AoA)` possible items (cognates = `r nrow(CognateList_nouns_AoA[which(CognateList_nouns_AoA$cognate_status=="cognate"),])`, non-cognates = `r nrow(CognateList_nouns_AoA[which(CognateList_nouns_AoA$cognate_status=="non-cognate"),])`). Using the optmatch package (Version 0.9.14; Hansen & Klopfer, 2006) in the R statistical language (R Core Team, 2019), each cognate item was matched to a non-cognate item according to the typical age of acquisition in both English and French obtained from the wordbankr package (Version 0.3.1; Braginsky, 2018) with the closest match possible on word category. There were `r nrow(CognateList_matching[which(CognateList_matching$exact_match=="exact"),])/2` pairs that matched exactly based on these criteria. For example, the cognate pair "chair"–_"chaise"_ and the non-cognate pair "bed"–_"lit"_ matched because they are typically acquired at age 21 months in English and French and are both in the furniture category (Frank et al., 2017). The remaining `r nrow(CognateList_matching[which(CognateList_matching$exact_match=="non-exact"),])/2` pairs were matched on age of acquisition as well, allowing a possible one-month deviation in either English, French or both. For example, the cognate pair "mittens"–_"mitaine"_ and the non-cognate pair "slipper"–_"pantoufle"_ matched since the English words are acquired at 28 and 27 months respectively (one-month deviation), both French words are acquired at 22 months of age (Frank et al., 2017), and both are clothing. Thus, the final items (`r nrow(CognateList_matching[which(CognateList_matching$cognate_status=="cognate"),])` cognates, `r nrow(CognateList_matching[which(CognateList_matching$cognate_status=="non-cognate"),])` non-cognates) included in the matched list were as similar as possible in all respects except their cognate status. 

## Analytical Strategy
Analyses were run on two different dependent variables to examine whether bilingual infants would produce more cognates than non-cognates over their vocabulary development. The first dependent variable was the proportion of items on the word list that infants produced, where translation equivalents were counted as separate items. For example, the word “banana” would be counted as a produced cognate, whether or not its translation equivalent _“banane”_ was produced. The second dependent variable was the proportion of translation equivalent pairs infants produced. Here, pairs were counted only if the infant produced both items in a pair. For example, the pair “banana”–_“banane”_ was counted as a produced cognate pair if and only if the child could produce both words in the pair.

For each dependent variable we conducted analyses using (1) the complete list of cognates and non-cognates (`r nrow(CognateList_full_clean)` translation equivalents pairs in total) and then restricted the analysis to (2) a matched list (nouns only and matched on age of acquisition; `r nrow(CognateList_matching)` translation equivalent pairs in total). Based on the two dependent variables and the two sets of words, we therefore ran a total of four models. Logistic mixed effects analyses were performed in the R statistical language (Version 4.0.2; R Core Team, 2019) using the lme4 package (Bates et al., 2015). Regression weights reflected the total number of cognates and non-cognates to account for the different number of words between the cognate and non-cognate lists. The lmerTest package (Kuznetsova et al., 2017) was used to calculate p-values. Goodness-of-fit tests for the logistic regression models were estimated using the DHARMa package (Hartig, 2022). Analysis scripts and the data set used in the present study are available at [https://osf.io/rh7av/].

# Results
## Descriptive Measures of Number of Words Produced
```{r include=FALSE, cache=TRUE}
keepers_cognate_full <- read.csv(here::here("data_keepers/public_keepers_cognate_full.csv"))

# calculate to see: (1) how many total words were produced; (2) if the TE pairs have been produced (i.e., both words in English and French are acquired); (3) how many cognate or non-cognate words have been produced (i.e., either word acquired only in one language or words acquired in both language)
keepers_cognate_full_wide <- keepers_cognate_full %>%
  # remove unnecessary variables
  select(-c(english_item, french_item)) %>%
  # calculate total number of eng/fr words acquired
  group_by(subject_id, n_months) %>%
  mutate(total_eng_acquired = sum(eng_acquired, na.rm=T),
         total_fr_acquired = sum(fr_acquired, na.rm=T)) %>%
  ungroup() %>%
  # calculate total number of words acquired (eng + fr)
  mutate(total_acquired = total_eng_acquired + total_fr_acquired,
         # percentage of words acquired
         percent_word_acquired = total_acquired/(537*2)*100) %>%
  # calculate to see if the TE pairs have been produced
  mutate(TE_acquired = case_when(eng_acquired == 1 & fr_acquired == 1 ~ 1,
                                 TRUE ~ 0)) %>%
  # remove the eng_acquired and fr_acquired variables
  select(-c(eng_acquired, fr_acquired)) %>%
  # unite french_item_id with word_pairs to facilitate the pivot_wider step
  unite(word_pairs_withID, c(english_item_id, word_pairs, french_item_id), remove=FALSE) %>%
  select(-c(word_pairs, english_item_id, french_item_id)) %>%
  # create wide data
  pivot_wider(names_from = word_pairs_withID, values_from = TE_acquired) %>%
  mutate(total_TE_acquired = rowSums(.[17:553], na.rm=T),
         percent_TE_acquired = case_when(cognate_status == "cognate" ~ total_TE_acquired/131*100,
                                       cognate_status == "non-cognate" ~ total_TE_acquired/406*100)) %>%
  # calculate total number of TEs (regardless of cognate status)
  group_by(subject_id, n_months) %>%
  mutate(total_all_TE_acquired = sum(total_TE_acquired, na.rm=T),
         percent_all_TE_acquired = total_all_TE_acquired/537*100) %>%
  ungroup()

# descriptives
## mean total words acquired
full_mean_total <- keepers_cognate_full_wide %>%
  select(subject_id, n_months, total_eng_acquired, total_fr_acquired, total_acquired, percent_word_acquired, 
         total_all_TE_acquired, percent_all_TE_acquired) %>%
  distinct(subject_id, n_months, .keep_all = TRUE) %>%
  pivot_longer(-c(subject_id, n_months), names_to = "type", values_to = "number") %>%
  group_by(type) %>%
  summarize(n_subject = n_distinct(subject_id),
            mean_number = mean(number, na.rm=T),
            sd_number = sd(number, na.rm=T),
            min_number = min(number, na.rm=T),
            max_number = max(number, na.rm=T))
```
Out of the complete list (a possible `r nrow(CognateList_full_clean)` translation equivalent pairs with `r nrow(CognateList_full_clean)` × 2 = `r nrow(CognateList_full_clean)*2` words), bilingual infants on average produced a total of `r round(full_mean_total[which(full_mean_total$type=="total_acquired"),3],0)` words (SD = `r round(full_mean_total[which(full_mean_total$type=="total_acquired"),4],2)`), with a range of `r round(full_mean_total[which(full_mean_total$type=="total_acquired"),5],0)` – `r round(full_mean_total[which(full_mean_total$type=="total_acquired"),6],0)` words, which constituted `r round(full_mean_total[which(full_mean_total$type=="percent_word_acquired"),3],1)`% of the words on the complete list. Moreover, bilingual infants produced an average of `r round(full_mean_total[which(full_mean_total$type=="total_all_TE_acquired"),3],0)` complete translation equivalent pairs where both the English and French words were produced (SD = `r round(full_mean_total[which(full_mean_total$type=="total_all_TE_acquired"),4],2)`, range = `r round(full_mean_total[which(full_mean_total$type=="total_all_TE_acquired"),5],0)` – `r round(full_mean_total[which(full_mean_total$type=="total_all_TE_acquired"),6],0)`), which constituted `r round(full_mean_total[which(full_mean_total$type=="percent_all_TE_acquired"),3],1)`% of the translation equivalent pairs on the complete list.

```{r include=FALSE, cache=TRUE}
keepers_cognate_matched <- read.csv(here::here("data_keepers/public_keepers_cognate_matched.csv"))

# calculate to see: (1) how many total words were produced; (2) if the TE pairs have been produced (i.e., both words in English and French are acquired)
# (3) how many cognate or non-cognate words have been produced (i.e., either word acquired only in one language or words acquired in both language)
keepers_cognate_matched_wide <- keepers_cognate_matched %>%
  # remove unnecessary variables
  select(-c(english_item_id, english_item, Eng_AoA, french_item_id, french_item, Fr_AoA)) %>%
  # calculate total number of eng/fr words acquired within the matched list
  group_by(subject_id, n_months) %>%
  mutate(total_eng_acquired = sum(eng_acquired, na.rm=T),
         total_fr_acquired = sum(fr_acquired, na.rm=T)) %>%
  ungroup() %>%
  # calculate total number of words acquired (eng + fr)
  mutate(total_acquired = total_eng_acquired + total_fr_acquired,
         # percentage of words acquired
         percent_word_acquired = total_acquired/(162*2)*100) %>%
  # calculate to see if the TE pairs have been produced
  mutate(TE_acquired = case_when(eng_acquired == 1 & fr_acquired == 1 ~ 1,
                                 TRUE ~ 0))  %>%
  # remove the eng_acquired and fr_acquired variables
  select(-c(eng_acquired, fr_acquired)) %>%
  pivot_wider(names_from = word_pairs, values_from = TE_acquired) %>%
  mutate(total_TE_acquired = rowSums(.[17:178], na.rm=T)) %>%
  mutate(percent_TE_acquired = total_TE_acquired/81*100) %>%
  # calculate total number of TEs (regardless of cognate status)
  group_by(subject_id, n_months) %>%
  mutate(total_all_TE_acquired = sum(total_TE_acquired, na.rm=T),
         percent_all_TE_acquired = total_all_TE_acquired/162*100) %>%
  ungroup()

# descriptives
## mean total words acquired
matched_mean_total <- keepers_cognate_matched_wide %>%
  select(subject_id, n_months, total_eng_acquired, total_fr_acquired, total_acquired, percent_word_acquired, 
         total_all_TE_acquired, percent_all_TE_acquired) %>%
  distinct(subject_id, n_months, .keep_all = TRUE) %>%
  pivot_longer(-c(subject_id, n_months), names_to = "type", values_to = "number") %>%
  group_by(type) %>%
  summarize(n_subject = n_distinct(subject_id),
            mean_number = mean(number, na.rm=T),
            sd_number = sd(number, na.rm=T),
            min_number = min(number, na.rm=T),
            max_number = max(number, na.rm=T))
```
Restricting to the matched list which contained `r nrow(CognateList_matching)` translation equivalent pairs with `r nrow(CognateList_matching)` × 2 = `r nrow(CognateList_matching)*2` words, bilingual infants produced an average of `r round(matched_mean_total[which(matched_mean_total$type=="total_acquired"),3],0)` words (SD = `r round(matched_mean_total[which(matched_mean_total$type=="total_acquired"),4],2)`, range = `r round(matched_mean_total[which(matched_mean_total$type=="total_acquired"),5],0)` – `r round(matched_mean_total[which(matched_mean_total$type=="total_acquired"),6],0)`), which constituted `r round(matched_mean_total[which(matched_mean_total$type=="percent_word_acquired"),3],1)`% of the words on the matched list. On average, bilingual infants produced a total of `r round(matched_mean_total[which(matched_mean_total$type=="total_all_TE_acquired"),3],0)` complete translation equivalent pairs (SD = `r round(matched_mean_total[which(matched_mean_total$type=="total_all_TE_acquired"),4],2)`, range = `r round(matched_mean_total[which(matched_mean_total$type=="total_all_TE_acquired"),5],0)` – `r round(matched_mean_total[which(matched_mean_total$type=="total_all_TE_acquired"),6],0)`), which constituted `r round(matched_mean_total[which(matched_mean_total$type=="percent_all_TE_acquired"),3],1)`% of translation equivalent pairs on the matched list.  

## Dependent Variable 1: Cognate Words Versus Non-Cognate Words
```{r include=FALSE, cache=TRUE}
# average age in the sample
mean_age <- keepers_cognate_full %>% 
  distinct(subject_id, .keep_all = T) %>% 
  summarize(mean_age = mean(age_days, na.rm = T))
```
In this analysis, the dependent variable was the total proportion of words infants produced on the relevant list. Proportion was used as opposed to raw number of words to provide a more comparable description of production of cognates versus non-cognates, since the number of cognate words and non-cognates words differed especially in the complete list. Our predictor variables were age (in days) and cognate status. Age was continuous and was centered at the mean age of `r round(mean_age,1)` days (approximately `r round(mean_age/30,0)` months) for ease of interpretation. Cognate status was categorical with two levels (cognates versus non-cognates) with non-cognates as the reference level. We ran separate logistic regression models for the complete and matched lists. The initial model specification included a random slope of age and cognate status by participants, which was pruned to a random intercept to achieve model convergence. The final model was:

proportion_word ~ age * cognate_status + (1|participant)

### Complete List
```{r include=FALSE, cache=TRUE}
# calculate to see how many cognate or non-cognate words have been produced (i.e., either word acquired only in one language or words acquired in both language)
keepers_cognate_full_notPair <- keepers_cognate_full %>%
  # remove unnecessary variables
  select(-c(english_item_id, english_item, french_item_id, french_item)) %>%
  # calculate total number of eng/fr words acquired within the full list
  group_by(subject_id, n_months, age_days, cognate_status) %>%
  summarize(n_Eng_acquired = sum(eng_acquired, na.rm=T),
            n_Fr_acquired = sum(fr_acquired, na.rm=T))%>%
  # calculate total number of words acquired within the full list
  mutate(n_word_acquired = n_Eng_acquired + n_Fr_acquired,
         # percentage of words acquired within the full list
         percent_word_acquired = case_when(cognate_status == "cognate" ~ n_word_acquired/(131*2)*100,
                                       cognate_status == "non-cognate" ~ n_word_acquired/(406*2)*100),
         # proportion of words acquired within the full list
         prop_word_acquired = case_when(cognate_status == "cognate" ~ n_word_acquired/(131*2),
                                       cognate_status == "non-cognate" ~ n_word_acquired/(406*2))) %>%
  ungroup() 

# descriptive
## mean number of cognates/non-cognates acquired
full_mean_cognates <- keepers_cognate_full_notPair %>%
  group_by(cognate_status) %>%
  summarize(mean_n_word = mean(n_word_acquired, na.rm=T),
            sd_n_word = sd(n_word_acquired, na.rm=T),
            min_n_word = min(n_word_acquired, na.rm=T),
            max_n_word = max(n_word_acquired, na.rm=T))

full_mean_cognates_prop <- keepers_cognate_full_notPair %>%
  group_by(cognate_status) %>%
  summarize(mean_prop_word = mean(prop_word_acquired, na.rm=T),
            sd_prop_word = sd(prop_word_acquired, na.rm=T),
            min_prop_word = min(prop_word_acquired, na.rm=T),
            max_prop_word = max(prop_word_acquired, na.rm=T))

# logistic mixed-effects model
Analysis1_full_logistic_model <- keepers_cognate_full_notPair %>%
  # scale and center age for ease of interpretation
  mutate(age_days = scale(age_days, scale = FALSE, center = TRUE)) %>%
  # reverse code "cognate_status"
  mutate(cognate_status = fct_rev(cognate_status)) %>%
  # create a weight variable for the model
  mutate(total_number = case_when(cognate_status == "cognate" ~ 131*2,
                        cognate_status == "non-cognate" ~ 406*2)) %>%
  # the glmer model
  glmer(prop_word_acquired ~ cognate_status * age_days + (1|subject_id),
        family = binomial, weights = total_number,
        data = .,
        nAGQ = 0)

Analysis1_full_logistic_model_coefs <- summary(Analysis1_full_logistic_model)$coef %>%
  as.data.frame %>%
  mutate_at(c("Estimate","Std. Error", "z value", "Pr(>|z|)"), 
            function (x) signif(x, digits = 3)) %>% 
  rename(SE = `Std. Error`, 
         z = `z value`,
         p = `Pr(>|z|)`) %>%
  mutate_if(is.numeric, round, digits = 3) %>%
  mutate(p = case_when(
    p < .001 ~ "<.001",
    p < .01 ~ "<.01",
    p < .05 ~ "<.05",
    TRUE ~ as.character(p)))

rownames(Analysis1_full_logistic_model_coefs) <- c("Intercept", "cognate_status", "age_days", 
                     "cognate_status * age_days") 

# goodness of fit
library(DHARMa)
testUniformity(Analysis1_full_logistic_model)
testDispersion(Analysis1_full_logistic_model) 
plotQQunif(Analysis1_full_logistic_model)


# logistic mixed-effects model 2 (counts as DV instead of prop) returns the same result
check <- keepers_cognate_full %>%
  # remove unnecessary variables
  select(-c(english_item, french_item, english_item_id, french_item_id, matches("total_"), matches("completed_"))) %>%
  pivot_longer(c(eng_acquired, fr_acquired), names_to = "language", values_to = "acquired") %>%
  # scale and center age for ease of interpretation
  mutate(age_days = scale(age_days, scale = FALSE, center = TRUE)) %>%
  # reverse code "cognate_status"
  mutate(cognate_status = fct_rev(cognate_status)) %>%
  # the glmer model
  glmer(acquired ~ cognate_status * age_days + (1|subject_id),
        family = binomial,
        data = ., 
        nAGQ = 0)

```
Out of the complete list which contained 262 cognate words (i.e., adding the 131 English cognate words and 131 French cognate words) and 812 non-cognate words (i.e., adding the 406 English non-cognate words and 406 French non-cognate words), bilingual infants produced an average of `r round(full_mean_cognates[which(full_mean_cognates$cognate_status=="cognate"),2],0)` cognate words (SD = `r round(full_mean_cognates[which(full_mean_cognates$cognate_status=="cognate"),3],2)`, range = `r round(full_mean_cognates[which(full_mean_cognates$cognate_status=="cognate"),4],0)` – `r round(full_mean_cognates[which(full_mean_cognates$cognate_status=="cognate"),5],0)`) and `r round(full_mean_cognates[which(full_mean_cognates$cognate_status=="non-cognate"),2],0)` non-cognate words (SD = `r round(full_mean_cognates[which(full_mean_cognates$cognate_status=="non-cognate"),3],2)`, range = `r round(full_mean_cognates[which(full_mean_cognates$cognate_status=="non-cognate"),4],0)` – `r round(full_mean_cognates[which(full_mean_cognates$cognate_status=="non-cognate"),5],0)`). The proportion of cognate words produced was `r round(full_mean_cognates_prop[which(full_mean_cognates_prop$cognate_status=="cognate"),2],2)` (SD = `r round(full_mean_cognates_prop[which(full_mean_cognates_prop$cognate_status=="cognate"),3],2)`, range = `r round(full_mean_cognates_prop[which(full_mean_cognates_prop$cognate_status=="cognate"),4],2)` – `r round(full_mean_cognates_prop[which(full_mean_cognates_prop$cognate_status=="cognate"),5],2)`), whereas the proportion of non-cognate words produced was `r round(full_mean_cognates_prop[which(full_mean_cognates_prop$cognate_status=="non-cognate"),2],2)` (SD = `r round(full_mean_cognates_prop[which(full_mean_cognates_prop$cognate_status=="non-cognate"),3],2)`, range = `r round(full_mean_cognates_prop[which(full_mean_cognates_prop$cognate_status=="non-cognate"),4],2)` – `r round(full_mean_cognates_prop[which(full_mean_cognates_prop$cognate_status=="non-cognate"),5],2)`). A Q-Q plot visualization and goodness-of-fit tests on the model’s residuals showed that our model had a good model fit, _D_ = `r round(testUniformity(Analysis1_full_logistic_model)$statistic,2)`, $p = `r papaja::printp(testUniformity(Analysis1_full_logistic_model)$p.value)`$. Table 1 shows the coefficient estimates for the model and Figure 1 Panel A visualizes the model. We observed significant main effects of age and cognate status, as well as a significant interaction. Overall, the pattern of results indicated that infants produced a greater proportion of cognates than non-cognates, with a slightly steeper learning curve for non-cognates than for cognates, although non-cognate production did not “catch up” to cognate production during the ages we observed.

### Matched List
```{r include=FALSE, cache=TRUE}
# calculate to see how many cognate or non-cognate words have been produced (i.e., either word acquired only in one language or words acquired in both language)
keepers_cognate_matched_notPair <- keepers_cognate_matched %>%
  # remove unnecessary variables
  select(-c(english_item_id, english_item, Eng_AoA, french_item_id, french_item, Fr_AoA)) %>%
  # calculate total number of eng/fr words acquired within the matched list
  group_by(subject_id, n_months, age_days, cognate_status) %>%
  summarize(n_Eng_acquired = sum(eng_acquired, na.rm=T),
            n_Fr_acquired = sum(fr_acquired, na.rm=T))%>%
  # calculate total number of words acquired within the matched list
  mutate(n_word_acquired = n_Eng_acquired + n_Fr_acquired,
         percent_word_acquired = n_word_acquired/(81*2)*100, # percentage of words acquired within the matched list
         prop_word_acquired = n_word_acquired/(81*2)) %>%
  ungroup() 

# descriptive
## mean number of cognates/non-cognates acquired
matched_mean_cognates <- keepers_cognate_matched_notPair %>%
  group_by(cognate_status) %>%
  summarize(mean_n_word = mean(n_word_acquired, na.rm=T),
            sd_n_word = sd(n_word_acquired, na.rm=T),
            min_n_word = min(n_word_acquired, na.rm=T),
            max_n_word = max(n_word_acquired, na.rm=T))

matched_mean_cognates_prop <- keepers_cognate_matched_notPair %>%
  group_by(cognate_status) %>%
  summarize(mean_prop_word = mean(prop_word_acquired, na.rm=T),
            sd_prop_word = sd(prop_word_acquired, na.rm=T),
            min_prop_word = min(prop_word_acquired, na.rm=T),
            max_prop_word = max(prop_word_acquired, na.rm=T))


# logistic mixed-effects model
Analysis1_matched_logistic_model <- keepers_cognate_matched_notPair %>%
  # scale and center age for ease of interpretation
  mutate(age_days = scale(age_days, scale = FALSE, center = TRUE)) %>%
  # reverse code "cognate_status"
  mutate(cognate_status = fct_rev(cognate_status)) %>%
  # create a weight variable for the model
  mutate(total_number = 81*2) %>%
  # the glmer model
  glmer(prop_word_acquired ~ cognate_status * age_days + (1|subject_id),
        family = binomial, weights = total_number,
        data = .,
        nAGQ = 0)

Analysis1_matched_logistic_model_coefs <- summary(Analysis1_matched_logistic_model)$coef %>%
  as.data.frame %>%
  mutate_at(c("Estimate","Std. Error", "z value", "Pr(>|z|)"), 
            function (x) signif(x, digits = 3)) %>% 
  rename(SE = `Std. Error`, 
         z = `z value`,
         p = `Pr(>|z|)`) %>%
  mutate_if(is.numeric, round, digits = 3) %>%
  mutate(p = case_when(
    p < .001 ~ "<.001",
    p < .01 ~ "<.01",
    p < .05 ~ "<.05",
    TRUE ~ as.character(p)))

rownames(Analysis1_matched_logistic_model_coefs) <- c("Intercept", "cognate_status", "age_days", 
                     "cognate_status * age_days") 


# goodness of fit
library(DHARMa)
testUniformity(Analysis1_matched_logistic_model)
testDispersion(Analysis1_matched_logistic_model) 
plotQQunif(Analysis1_matched_logistic_model)

```

Out of the 162 cognate (i.e., adding the 81 English cognate words and 81 French cognate words) and 162 non-cognate words (i.e., adding the 81 English non-cognate words and 81 French non-cognate words) on the matched list, bilingual infants produced an average of `r round(matched_mean_cognates[which(matched_mean_cognates$cognate_status=="cognate"),2],0)` cognate words (SD = `r round(matched_mean_cognates[which(matched_mean_cognates$cognate_status=="cognate"),3],2)`, range = `r round(matched_mean_cognates[which(matched_mean_cognates$cognate_status=="cognate"),4],0)` – `r round(matched_mean_cognates[which(matched_mean_cognates$cognate_status=="cognate"),5],0)`) and `r round(matched_mean_cognates[which(matched_mean_cognates$cognate_status=="non-cognate"),2],0)` non-cognate words (SD = `r round(matched_mean_cognates[which(matched_mean_cognates$cognate_status=="non-cognate"),3],2)`, range = `r round(matched_mean_cognates[which(matched_mean_cognates$cognate_status=="non-cognate"),4],0)` – `r round(matched_mean_cognates[which(matched_mean_cognates$cognate_status=="non-cognate"),5],0)`). The overall mean proportion of cognate words produced was `r round(matched_mean_cognates_prop[which(matched_mean_cognates_prop$cognate_status=="cognate"),2],2)` of words (SD = `r round(matched_mean_cognates_prop[which(matched_mean_cognates_prop$cognate_status=="cognate"),3],2)`, range = `r round(matched_mean_cognates_prop[which(matched_mean_cognates_prop$cognate_status=="cognate"),4],2)` – `r round(matched_mean_cognates_prop[which(matched_mean_cognates_prop$cognate_status=="cognate"),5],2)`), whereas the proportion of non-cognate words produced was `r round(matched_mean_cognates_prop[which(matched_mean_cognates_prop$cognate_status=="non-cognate"),2],2)` (SD = `r round(matched_mean_cognates_prop[which(matched_mean_cognates_prop$cognate_status=="non-cognate"),3],2)`, range = `r round(matched_mean_cognates_prop[which(matched_mean_cognates_prop$cognate_status=="non-cognate"),4],2)` – `r round(matched_mean_cognates_prop[which(matched_mean_cognates_prop$cognate_status=="non-cognate"),5],2)`). A Q-Q plot visualization as well as goodness-of-fit tests on the model’s residuals showed a good model fit, _D_ = `r round(testUniformity(Analysis1_matched_logistic_model)$statistic,2)`, $p = `r papaja::printp(testUniformity(Analysis1_matched_logistic_model)$p.value)`$. Table 1 also shows the coefficient estimates for the matched list model and Figure 1 Panel B visualizes the model. Similar to the patterns reported in the complete list model, there were significant effects of age and cognate status, once again showing that infants produced a greater proportion of cognates than non-cognates on the matched list. However, for the matched list there was no interaction between cognate status and age, indicating that the magnitude of the cognate advantage for this list was stable as infants grew older. 

```{r Fig1, fig.cap="Proportion of words produced by age and cognate status, with Panel A representing the complete list and Panel B representing the matched list. Note that the black dashed line represents the mean age of 547.6 days which serves as the reference level for age in our models.", echo=FALSE, dpi=600, fig.align='center', fig.height=4, out.width="120%", cache = TRUE}

prop_word_full <- keepers_cognate_full_notPair %>%
  select(subject_id, n_months, age_days, cognate_status, prop_word_acquired) %>%
  rename(full_prop_word = prop_word_acquired) 

prop_word_matched <- keepers_cognate_matched_notPair %>%
  select(subject_id, n_months, age_days, cognate_status, prop_word_acquired) %>%
  rename(matched_prop_word = prop_word_acquired)

Figure1 <- right_join(prop_word_full, prop_word_matched) %>%
  pivot_longer(c(full_prop_word, matched_prop_word), names_to = "list", values_to = "prop_word") %>%
  mutate(list = recode(list, 
                        full_prop_word = "Panel A: Complete list",
                        matched_prop_word = "Panel B: Matched list"),
         cognate_status = recode(cognate_status,
                                  "cognate" = "Cognate",
                                  "non-cognate" = "Non-cognate")) %>%
  group_by(subject_id, n_months, list, cognate_status) %>%
  ggplot(aes(x = age_days, y = prop_word, color = cognate_status)) +
  facet_wrap(. ~ list) +
  geom_smooth(method = "glm", se = FALSE, method.args = list(family = binomial)) + 
  geom_point(stat = "identity", position = "identity", shape = 23, alpha = 0.3) +
  geom_vline(xintercept = 547.6, linetype="dashed", alpha = 0.5) +
  labs(x = "\n Age (in days)", 
       y = "Proportion of words produced",
       color = "Cognate status") + 
  ylim(0, 1) +
  theme_minimal() +
  theme(text = element_text(size=12),
        strip.text.x = element_text(size = 12),
        legend.position = "bottom",
        panel.spacing = unit(1.5, "lines")) 

Figure1

ggsave("Figure1.png", Figure1,
       width = 9, height = 6)
```

```{r Table 1, results = "asis", cache = TRUE}
Table1 <- cbind(Analysis1_full_logistic_model_coefs, Analysis1_matched_logistic_model_coefs)

papaja::apa_table(Table1, format.args = list(digits = 3),
                  col.names =c("","Estimate", "$SE$","$z$","$p$","Estimate", "$SE$","$z$","$p$"),
                  align=c("l","c","c","c","c","c","c","c", "c"),
                  col_spanners = list("Complete list" = c(2,5), "Matched list" = c(6,9)),
                  caption="Table 1. Coefficient estimates from the mixed-effects logistic models predicting proportion of words produced.",
                  landscape = FALSE)
```

## Dependent Variable 2: Cognate Pairs Versus Non-Cognate Pairs
In this analysis, the proportion of translation equivalent pairs produced was entered as the dependent variable. Age and cognate status were entered as our predictor variables, with non-cognates set as the reference level. Again, we ran separate logistic models for the complete and matched lists. The initial model specification, which included a random effect of age and cognate status by participants, had to be reduced for model convergence; therefore, the final model was:

proportion_pair ~ age * cognate_status + (1|participant)

### Complete List
```{r include=FALSE, cache=TRUE}
# mean number of cognates/non-cognates acquired
## total number
full_mean_total_TE <- keepers_cognate_full_wide %>%
  group_by(cognate_status) %>%
  summarize(mean_TE = mean(total_TE_acquired, na.rm=T),
            sd_TE = sd(total_TE_acquired, na.rm=T),
            min_TE = min(total_TE_acquired, na.rm=T),
            max_TE = max(total_TE_acquired, na.rm=T))

## proportion
full_mean_prop_TE <- keepers_cognate_full_wide %>%
  mutate(prop_TE_acquired = percent_TE_acquired/100) %>%
  group_by(cognate_status) %>%
  summarize(mean_TE = mean(prop_TE_acquired, na.rm=T),
            sd_TE = sd(prop_TE_acquired, na.rm=T),
            min_TE = min(prop_TE_acquired, na.rm=T),
            max_TE = max(prop_TE_acquired, na.rm=T))


# logistic model
Analysis2_full_logistic_model <- keepers_cognate_full_wide %>%
  # reverse-order cognate status, so that the non-cognate status would be the reference level
  mutate(cognate_status = fct_rev(cognate_status)) %>%
  # scale and center age for ease of interpretation
  mutate(age_days = scale(age_days, scale = FALSE, center = TRUE)) %>%
  # change percentage to proportion
  mutate(prop_TE_acquired = percent_TE_acquired/100) %>%
  # create a weight variable for the model
  mutate(total_number = case_when(cognate_status == "cognate" ~ 131,
                        cognate_status == "non-cognate" ~ 406)) %>%
  # the glmer model
  glmer(prop_TE_acquired ~ cognate_status * age_days + (1|subject_id),
        family = binomial, weights = total_number,
        data = .,
        nAGQ = 0)


Analysis2_full_logistic_model_coefs <- summary(Analysis2_full_logistic_model)$coef %>%
  as.data.frame %>%
  mutate_at(c("Estimate","Std. Error", "z value", "Pr(>|z|)"), 
            function (x) signif(x, digits = 3)) %>% 
  rename(SE = `Std. Error`, 
         z = `z value`,
         p = `Pr(>|z|)`) %>%
  mutate_if(is.numeric, round, digits = 3) %>%
  mutate(p = case_when(
    p < .001 ~ "<.001",
    p < .01 ~ "<.01",
    p < .05 ~ "<.05",
    TRUE ~ as.character(p)))

rownames(Analysis2_full_logistic_model_coefs) <- c("Intercept", "cognate_status", "age_days", 
                     "cognate_status * age_days") 


# goodness of fit
library(DHARMa)
testUniformity(Analysis2_full_logistic_model)
testDispersion(Analysis2_full_logistic_model) 
plotQQunif(Analysis2_full_logistic_model)

```

Out of the complete list which contained `r nrow(CognateList_full_clean)` translation equivalent pairs (`r nrow(CognateList_full_clean[which(CognateList_full_clean$cognate_status=="cognate"),])` cognates and `r nrow(CognateList_full_clean[which(CognateList_full_clean$cognate_status=="non-cognate"),])` non-cognates), infants produced an average of `r round(full_mean_total_TE[which(full_mean_total_TE$cognate_status=="cognate"),2],0)` cognate pairs (SD = `r round(full_mean_total_TE[which(full_mean_total_TE$cognate_status=="cognate"),3],2)`, range = `r round(full_mean_total_TE[which(full_mean_total_TE$cognate_status=="cognate"),4],0)` – `r round(full_mean_total_TE[which(full_mean_total_TE$cognate_status=="cognate"),5],0)`) and `r round(full_mean_total_TE[which(full_mean_total_TE$cognate_status=="non-cognate"),2],0)` non-cognate pairs (SD = `r round(full_mean_total_TE[which(full_mean_total_TE$cognate_status=="non-cognate"),3],2)`, range = `r round(full_mean_total_TE[which(full_mean_total_TE$cognate_status=="non-cognate"),4],0)` – `r round(full_mean_total_TE[which(full_mean_total_TE$cognate_status=="non-cognate"),5],0)`). The proportion of cognate pairs produced was `r round(full_mean_prop_TE[which(full_mean_prop_TE$cognate_status=="cognate"),2],2)` (SD = `r round(full_mean_prop_TE[which(full_mean_prop_TE$cognate_status=="cognate"),3],2)`, range = `r round(full_mean_prop_TE[which(full_mean_prop_TE$cognate_status=="cognate"),4],2)` – `r round(full_mean_prop_TE[which(full_mean_prop_TE$cognate_status=="cognate"),5],2)`) whereas the proportion of non-cognate pairs produced was `r round(full_mean_prop_TE[which(full_mean_prop_TE$cognate_status=="non-cognate"),2],2)` (SD = `r round(full_mean_prop_TE[which(full_mean_prop_TE$cognate_status=="non-cognate"),3],2)`, range = `r round(full_mean_prop_TE[which(full_mean_prop_TE$cognate_status=="non-cognate"),4],2)` – `r round(full_mean_prop_TE[which(full_mean_prop_TE$cognate_status=="non-cognate"),5],2)`). A Q-Q plot visualization and goodness-of-fit tests on the model’s residuals revealed that our model showed a good model fit, _D_ = `r round(testUniformity(Analysis2_full_logistic_model)$statistic,2)`, $p = `r papaja::printp(testUniformity(Analysis2_full_logistic_model)$p.value)`$. Table 2 shows the coefficient estimates for the model and Figure 2 Panel A visualizes the model. There were significant effects of age and cognate status, showing that overall infants produced a greater proportion of cognates than non-cognates. Similar to the pattern reported in the first set of analyses, the interaction between age and cognate status suggested a slightly steeper learning curve for non-cognates than cognates, although an advantage for cognates was still apparent even at 27 months.

### Matched List
```{r include=FALSE, cache=TRUE}
# mean number of cognates/non-cognates acquired
## total number
matched_mean_total_TE <- keepers_cognate_matched_wide %>%
  group_by(cognate_status) %>%
  summarize(mean_TE = mean(total_TE_acquired, na.rm=T),
            sd_TE = sd(total_TE_acquired, na.rm=T),
            min_TE = min(total_TE_acquired, na.rm=T),
            max_TE = max(total_TE_acquired, na.rm=T))

## proportion
matched_mean_prop_TE <- keepers_cognate_matched_wide %>%
  mutate(prop_TE_acquired = percent_TE_acquired/100) %>%
  group_by(cognate_status) %>%
  summarize(mean_TE = mean(prop_TE_acquired, na.rm=T),
            sd_TE = sd(prop_TE_acquired, na.rm=T),
            min_TE = min(prop_TE_acquired, na.rm=T),
            max_TE = max(prop_TE_acquired, na.rm=T))


# logistic model
Analysis2_matched_logistic_model <- keepers_cognate_matched_wide %>%
  # reverse-order cognate status, so that the non-cognate status would be the reference level
  mutate(cognate_status = fct_rev(cognate_status)) %>%
  # scale and center age for ease of interpretation
  mutate(age_days = scale(age_days, scale = FALSE, center = TRUE)) %>%
  # change percentage to proportion
  mutate(prop_TE_acquired = percent_TE_acquired/100) %>%
  # create a weight variable for the model
  mutate(total_number = 81) %>%
  # the glmer model
  glmer(prop_TE_acquired ~ cognate_status * age_days + (1|subject_id),
        family = binomial, weights = total_number,
        data = .,
        nAGQ = 0)


Analysis2_matched_logistic_model_coefs <- summary(Analysis2_matched_logistic_model)$coef %>%
  as.data.frame %>%
  mutate_at(c("Estimate","Std. Error", "z value", "Pr(>|z|)"), 
            function (x) signif(x, digits = 3)) %>% 
  rename(SE = `Std. Error`, 
         z = `z value`,
         p = `Pr(>|z|)`) %>%
  mutate_if(is.numeric, round, digits = 3) %>%
  mutate(p = case_when(
    p < .001 ~ "<.001",
    p < .01 ~ "<.01",
    p < .05 ~ "<.05",
    TRUE ~ as.character(p)))

rownames(Analysis2_matched_logistic_model_coefs) <- c("Intercept", "cognate_status", "age_days", 
                     "cognate_status * age_days") 


# goodness of fit
library(DHARMa)
testUniformity(Analysis2_matched_logistic_model)
testDispersion(Analysis2_matched_logistic_model) 
plotQQunif(Analysis2_matched_logistic_model)

```

Out of the `r nrow(CognateList_matching)` translation equivalent pairs, bilingual infants produced an average of `r round(matched_mean_total_TE[matched_mean_total_TE$cognate_status=="cognate",2],0)` cognate pairs (SD = `r round(matched_mean_total_TE[matched_mean_total_TE$cognate_status=="cognate",3],2)`, range = `r round(matched_mean_total_TE[matched_mean_total_TE$cognate_status=="cognate",4],0)` – `r round(matched_mean_total_TE[matched_mean_total_TE$cognate_status=="cognate",5],0)`) and `r round(matched_mean_total_TE[matched_mean_total_TE$cognate_status=="non-cognate",2],0)` non-cognate pairs (SD = `r round(matched_mean_total_TE[matched_mean_total_TE$cognate_status=="non-cognate",3],2)`, range = `r round(matched_mean_total_TE[matched_mean_total_TE$cognate_status=="non-cognate",4],0)` – `r round(matched_mean_total_TE[matched_mean_total_TE$cognate_status=="non-cognate",5],0)`). The proportion of cognate pairs produced was `r round(matched_mean_prop_TE[which(matched_mean_prop_TE$cognate_status=="cognate"),2],2)` (SD = `r round(matched_mean_prop_TE[which(matched_mean_prop_TE$cognate_status=="cognate"),3],2)`, range = `r round(matched_mean_prop_TE[which(matched_mean_prop_TE$cognate_status=="cognate"),4],2)` – `r round(matched_mean_prop_TE[which(matched_mean_prop_TE$cognate_status=="cognate"),5],2)`) and the proportion of non-cognate pairs produced was `r round(matched_mean_prop_TE[which(matched_mean_prop_TE$cognate_status=="non-cognate"),2],2)`  (SD = `r round(matched_mean_prop_TE[which(matched_mean_prop_TE$cognate_status=="non-cognate"),3],2)`, range = `r round(matched_mean_prop_TE[which(matched_mean_prop_TE$cognate_status=="non-cognate"),4],2)` – `r round(matched_mean_prop_TE[which(matched_mean_prop_TE$cognate_status=="non-cognate"),5],2)`). A Q-Q plot visualization and the goodness-of-fit test on the model’s residuals (_D_ = `r round(testUniformity(Analysis2_matched_logistic_model)$statistic,2)`, $p = `r papaja::printp(testUniformity(Analysis2_matched_logistic_model)$p.value)`$) suggested that the logistic model did not fully capture the distribution of the data, but we nevertheless retained the model on theoretical grounds (the dependent variable was proportion) and to facilitate comparison to the previous models. The coefficient estimates for the matched list model is shown in Table 2, and Figure 2 Panel B visualizes the model. Similar to the results for the complete list, the main effects of age and cognate status were statistically significant, showing that infants produced a larger proportion of cognates than non-cognates. However, unlike the results for the complete list, the interaction between age and cognate status was not statistically significant, showing that the magnitude of the cognate difference was reasonably stable across age. 

```{r Fig2, fig.cap="Proportion of translation equivalent pairs produced by age and cognate status, with Panel A representing the complete list and Panel B representing the matched list. Note that the black dashed line represents the mean age of 547.6 days which serves as the reference level for age in our models.", echo=FALSE, dpi=600, fig.align='center', fig.height=4, out.width="120%", cache = TRUE}

prop_TE_full <- keepers_cognate_full_wide %>%
  select(subject_id, n_months, age_days, cognate_status, percent_TE_acquired) %>%
  mutate(full_prop_TE = percent_TE_acquired/100) %>%
  select(-(percent_TE_acquired))

prop_TE_matched <- keepers_cognate_matched_wide %>%
  select(subject_id, n_months, age_days, cognate_status, percent_TE_acquired) %>%
  mutate(matched_prop_TE = percent_TE_acquired/100) %>%
  select(-(percent_TE_acquired))

Figure2 <- right_join(prop_TE_full, prop_TE_matched) %>%
  pivot_longer(c(full_prop_TE, matched_prop_TE), names_to = "list", values_to = "prop_TE") %>%
  mutate(list = recode(list, 
                        full_prop_TE = "Panel A: Complete list",
                        matched_prop_TE = "Panel B: Matched list"),
         cognate_status = recode(cognate_status,
                                  "cognate" = "Cognate",
                                  "non-cognate" = "Non-cognate")) %>%
  group_by(subject_id, n_months, list, cognate_status) %>%
  ggplot(aes(x = age_days, y = prop_TE, color = cognate_status)) +
  facet_wrap(. ~ list) +
  geom_smooth(method = "glm", se = FALSE, method.args = list(family = binomial)) + 
  geom_point(stat = "identity", position = "identity", shape = 23, alpha = 0.3) +
  geom_vline(xintercept = 547.6, linetype = "dashed", alpha = 0.5) +
  labs(x = "\n Age (in days)", 
       y = "Proportion of translation equivalent pairs produced",
       color = "Cognate status") + 
  ylim(0,1) +
  theme_minimal() +
  theme(text = element_text(size=12),
        strip.text.x = element_text(size = 12),
        legend.position = "bottom",
        panel.spacing = unit(1.5, "lines")) 

Figure2

ggsave("Figure2.png", Figure2,
       width = 9, height = 6)
```

```{r Table 2, results = "asis", cache = TRUE}
Table2 <- cbind(Analysis2_full_logistic_model_coefs, Analysis2_matched_logistic_model_coefs)

papaja::apa_table(Table2, format.args = list(digits = 3),
                  col.names =c("","Estimate", "$SE$","$z$","$p$","Estimate", "$SE$","$z$","$p$"),
                  align=c("l","c","c","c","c","c","c","c", "c"),
                  col_spanners = list("Complete list" = c(2,5), "Matched list" = c(6,9)),
                  caption="Table 2. Coefficient estimates from the mixed-effects logistic models predicting proportion of translation equivalent pairs produced.",
                  landscape = FALSE)
```

## Summary of Analyses
Overall, the result patterns were largely consistent across the two sets of analyses, whereby bilingual infants produced a greater proportion of cognates than non-cognates. Infants increased their production of both cognates and non-cognates across age, and for the complete list (although not the matched list) this increase was slightly steeper for non-cognates than cognates, although production of cognates remained proportionally greater than that of cognates at the oldest age we observed (27 months). 

# Discussion
This current study evaluated whether phonological similarity facilitates vocabulary learning in bilinguals, by examining whether cognates are advantaged in bilingual infants’ early vocabulary production. Using monthly expressive vocabulary data, our longitudinal dataset revealed an overall advantage for cognates in infancy. Across ages, infants produced proportionally more cognates (e.g., English “banana” \textipa{/b@n\ae{}n@/}–French _”banane”_ \textipa{/b\textscripta{}n\textscripta{}n/}) than non-cognates (e.g., English “apple” \textipa{/\ae{}p@l/}–French _”pomme”_ \textipa{/p\textopeno{}m/}), although note that in raw terms children still produced a greater number of non-cognates than cognates, due to the greater absolute frequency of non-cognate translation equivalents on the French–English CDI checklists. 

Together with previous findings, our results begin to paint a developmental picture of the effects of cognate status on early vocabulary productions. Spanish and Catalan have both form-similar and form-identical cognates, and Bosch & Ramon-Casas (2014) reported a cognate advantage for form-identical but not form-similar cognates at 18 months. Cognates in French and English are almost exclusively form-similar, and we found an advantage for these form-similar in infants aged 16 to 27 months. Other studies have also reported an advantage for non-cognate translation equivalents (Bilson et al., 2015), which might vary with age (Tsui et al., in press). Overall, translation equivalents with the largest phonological overlap appear to be the most advantaged in early production and thus their effect might be detectable from age 18 months, with potential advantages in children’s production of form-similar and non-cognate translation equivalents strengthening across the second and third year of life. 

The robust cognate advantage across different bilingual infant populations points to the possibility that the origin of the cognate facilitation effect observed in childhood and in adulthood emerges from infancy. Previous studies which examined the cognate facilitation effect in bilingual adults and school-aged children have reported that bilinguals are better at processing cognates; for example, they can identify and/or name cognates more easily and quickly in a vocabulary task (Costa et al., 2000; Kelley & Kohnert, 2012; Sheng et al., 2016). Thus, the cognate facilitation effect appears to be robust in vocabulary production across the lifespan, with the advantage for cognates in production emerging early on, as our study results suggested. 

We interpret these results in light of theories that emphasize the interconnectedness of the two languages in the developing bilingual lexicon (DeAnda et al., 2016). Studies show that, even across languages, words that are semantically related are acquired sooner by bilingual children (Bilson et al., 2015) and are co-activated in language processing (e.g., DeAnda & Friend, 2020; Jardak & Byers-Heinlein, 2018; Singh, 2014). Moreover, young monolinguals find it easier to learn words that are phonologically similar to one another (Coady & Aslin, 2003; Demke et al., 2002; Jones & Brandt, 2019), and young bilinguals co-activate phonologically-related words both within and across languages (Von Holzen & Mani, 2012). These two sets of findings were confirmed in a study of monolingual children across 10 languages, who were more likely to acquire words with a high degree of semantic or phonological association (Fourtassi et al., 2020). Unique to bilinguals, cognates have a high degree of both semantic and phonological overlap, which our results show facilitate their acquisition.

There are several specific ways that cognates’ phonological and semantic overlap might advantage their learning. One possibility is that, for cognates, bilingual children might only need to map one phonological form (or slightly varied phonological forms for the cases of form-similar cognates) to label the same referent across the two languages, whereas for non-cognate translation equivalents bilingual children have to memorize two completely different forms for the same referent. It has been found that, indeed, bilingual children learning similar languages learn more cognate translation equivalents and have a larger vocabulary size in general (Gampe et al., 2021). Thus, transfer effects could explain the cognate advantage we observed in production, and would predict an early-emerging cognate advantage for word comprehension as well. Another possibility is that hearing a cognate word activates and strengthens phonological representations for both languages (e.g., hearing “banana” could activate and strengthen both “banana” and _“banane”_), thus accelerating cognate learning. Bilingual children have been found to identify and name cognates easier and faster than non-cognates, suggesting that the phonological overlap in cognates could support bilinguals’ lexical decoding and processing (Kelley & Kohnert, 2012; Sheng et al., 2016). Finally, the closer the phonological form of cognates, the more similar they might be for children to articulate. Support for such a hypothesis comes from study which showed that bilingual children not only learned phonologically-similar words faster but produced phonologically-similar nouns more frequently and more evenly than form-dissimilar nouns across their two languages (Schelletter, 2002). Note that these three possible mechanisms are not mutually exclusive, and could each contribute to the cognate advantage we observed.

There are several other factors that could also contribute to children’s faster learning of cognates than non-cognates. For example, Bosch & Ramon-Casas (2014) brought up several additional possibilities including frequency in the language input, reference to more complex concepts, or production difficulty due to changes in phonological forms, although they could not provide direct evidence due to the limited items on their vocabulary checklists. Our study attempted to account for several of these factors, by analyzing a subset of cognates and non-cognates that were carefully matched for part of speech, typical age of acquisition, and word category when possible. With this carefully controlled subset, we again found a production advantage for cognates. Thus, while such additional factors could potentially contribute to the cognate advantage, our results suggest that such third variable explanations are unlikely to underlie our results.

The cognate advantage can, at least in part, explain why bilingual children learning more similar languages show accelerated vocabulary development relative to bilinguals acquiring less similar languages (Blom et al., 2020; Gampe et al., 2021; Sheng et al., 2016). It has been shown that the more overlap shared across the two languages, the easier the words are learned by bilingual children (Bosma et al., 2019). Therefore, for those who are learning close language pairs that share a high degree of phonological overlap like Spanish and Catalan, their two languages share many cognates which sometimes are even form-identical, meaning that they are pronounced the same way in both languages (e.g, _“sí”_ /si/ meaning "yes" in both languages). On the other hand, for those who are learning languages that share a lesser degree of phonological similarity like English and French, there are potentially very few form-identical cognates. It is possible that when languages are very similar and share many form-identical cognates, bilingual infants can benefit from these words from a very young age. On the other hand, when languages are somewhat less similar and share mostly form-similar cognates, children may need more time to detect and benefit from cognates. Overall, we suggest that there is a gradual timeline for the facilitative effect of cognates in infancy, which starts off with form-identical cognates then form-similar cognates (Bosma et al., 2019). Future studies could include additional language pairs which are less similar than Spanish and Catalan but more similar than English and French, such as Spanish and Italian (Schepens et al., 2013), to directly compare the timeline regarding the acquisition of form-identical cognates, form-similar cognates, and non-cognates.

An important avenue for future research would be to examine whether the same cognate advantage would be observed in receptive vocabulary acquisition, and indeed some evidence points in this direction. Young bilingual infants show less perceptual sensitivity to cross-language phonological distinctions in cognates due to their phonological similarity (Ramon-Casas et al., 2009; Ramon-Casas & Bosch, 2010), suggesting that cognates may hold a different status in early bilinguals’ receptive lexicons compared to non-cognates. There is overall mixed evidence for whether the effect of cognate status is absent in comprehension (Schott & Byers-Heinlein, 2019) or is present in both comprehension and production (Kelley and Kohnert, 2012). However, it is possible that the cognate advantage is modulated by the level of difficulty of the vocabulary item for both comprehension and production. One study found that although the cognate advantage was observed in easier items, the effect was even greater in vocabulary items that were considered to be medium or hard (Kelley & Kohnert, 2012). This may suggest that infants would have a cognate advantage in any vocabulary task—either receptive or expressive, especially for less-familiar words where they may use the cognate word they have already acquired for help (Kelley & Kohnert, 2012), which is the case when infants are acquiring new words and learning to pronounce them. Therefore, we could expect a cognate advantage in both comprehension and production, serving different purposes: in comprehension, a cognate advantage would help activate the representations for the words in both languages, whereas in production, cognates may also facilitate the acquisition of the word in the individuals’ other language in terms of pronunciation, as was seen in our study. Future research could explore the difference between comprehension and production in bilingual infants’ language acquisition while simultaneously looking at the cognate advantage. 

# Conclusion
The present study demonstrated that French–English bilingual infants show an advantage for cognates in vocabulary production, with proportionally more cognates being produced than non-cognates. This finding can, at least in part, explain why children learning typologically similar languages show faster vocabulary growth than those learning more distant languages (Blom et al., 2020; Gampe et al., 2021; Sheng et al., 2016). Altogether, our study provides a greater understanding of the effect of similar-sounding words on infants’ language acquisition over time. Future studies with data from other populations of bilinguals will be important to more fully understand the effect of the cognate advantage in early bilingual vocabulary development. 

\newpage

# Acknowledgements
\noindent We are grateful to all the families who participated in this research. This manuscript derives from a thesis submitted by Lori Mitchell in partial fulfilment for an honours degree in Psychology at Concordia University. This work was supported by grants to Byers-Heinlein from the Natural Sciences and Engineering council of Canada [2018-04390]) and the National Institutes of Health [1R01HD095912-01A1]. Byers-Heinlein holds the Concordia University Research Chair in Bilingualism and Open Science. This study has been presented as a poster at the 2021 Boston University Conference on Language Development, and has been submitted as an abstract to the 2022 International Congress of Infant Studies. We thank the members of the Concordia Infant Research Lab for their comments on earlier versions of this paper.

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}

<div id = "refs"></div>
\endgroup
