---
title: "Supplemental Materials"
shorttitle: "Cognate advantage in bilingual infants"

header-includes:
  - \usepackage{amsmath}
  - \usepackage[labelformat=empty]{caption}
  - \usepackage{caption}
  - \usepackage[extra]{tipa}
  - \renewcommand{\topfraction}{1}
  - \renewcommand{\bottomfraction}{1}
  - \renewcommand{\textfraction}{.1}
  - \renewcommand{\floatpagefraction}{1}
  - \setcounter{topnumber}{9}
  - \setcounter{bottomnumber}{9}
  - \setcounter{totalnumber}{20}
  - \setcounter{dbltopnumber}{9}

author: 
  - name          : "Lori Mitchell"
    affiliation   : "1"
    email         : "lorimitchell7@hotmail.com"
  - name          : "Rachel Ka-Ying Tsui"
    affiliation   : "1" 
    email         : "rachelkytsui@gmail.com"
    corresponding : yes
    address       : "Department of Psychology, 7141 Sherbrooke St. West, Montreal, QC, Canada, H2T1V2. Rachel Ka-Ying Tsui is now at Laboratory for Language Development, RIKEN Center for Brain Science, 2-1 Hirosawa, Wako-shi, Saitama, Japan, 351-0198"
  - name          : "Krista Byers-Heinlein"
    affiliation   : "1" 
    email         : "k.byers@concordia.ca"
    address       : "Department of Psychology, 7141 Sherbrooke St. West, Montreal, QC, Canada, H2T1V2."


affiliation:
  - id            : "1"
    institution   : "Concordia University"

abstract: 

keywords          :

wordcount         : 

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no # "yes" to omit identifying information from the title page

#bibliography      : ["CogVocab_paper_references.bib"]
replace_ampersands: yes

class             : "man"
output            : papaja::apa6_pdf # change to apa6_docx for word document
citation_package  : biblatex
nocite: '@*' # include all the references in the .bib file, because in-text citations were mostly typed manually to fit the apa7 style
---

\captionsetup[table]{labelformat=empty}

```{r setup, include=FALSE, cache=TRUE}
library(rmarkdown)
library(knitr)
library(papaja)
library(kableExtra)
library(english)
library(linguisticsdown)
library(here)
library(tidyverse)
library(tidylog)
library(wordbankr)
library(matrixStats)
library(lme4)
library(lmerTest)
library(ggplot2)
library(patchwork)
library(gridExtra)
library("ggpubr")
library(broom)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)

```

```{r load_data, cache = TRUE}
# read in the final cleaned keepers_data
clean_demog <- read.csv(here::here("data_keepers/public_clean_demog.csv"))

exclusion <- read.csv(here::here("data_keepers/exclusions.csv"))

public_keepers_cognate_full <- read.csv(here::here("data_keepers/public_keepers_cognate_full.csv"))

public_keepers_cognate_matched <- read.csv(here::here("data_keepers/public_keepers_cognate_matched.csv"))
```

```{r include=FALSE, cache=TRUE}
# ---------- 25-75% language exposure criterion ----------
# keep only subjects with 25-75% language exposure
exceed_25_75_langexp <- clean_demog %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  mutate(flag_langexp = case_when(exposure_eng < 25 | exposure_eng > 75 ~ 1,
                                  exposure_fr < 25 | exposure_fr > 75 ~ 1,
                                  TRUE ~ 0)) %>%
  # which participant has a larger range throughout the study
  group_by(subject_id) %>%
  mutate(n_flag_langexp = sum(flag_langexp)) %>%
  filter(n_flag_langexp != 0) %>%
  # extract subject_id for later exclusion
  distinct(subject_id)

clean_demog_stringent <- clean_demog %>%
  anti_join(exceed_25_75_langexp)

# number of participants after removing these 9 participants
total_n_stringent <- clean_demog_stringent %>%
  filter(monolingual != 1) %>%
  distinct(subject_id) %>%
  nrow()

# number of girls
n_female_stringent <- clean_demog_stringent %>%
  filter(monolingual != 1) %>%
  select(subject_id, sex) %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  filter(sex == "Female") %>%
  nrow()

# Age range for 1st month data entry
mean_age_1st_entry_stringent <- clean_demog_stringent %>%
  filter(monolingual != 1) %>%
  filter(n_months == 1) %>%
  summarize(mean_age = mean(age_months, na.rm=T),
            sd_age = sd(age_months, na.rm=T),
            min_age = min(age_months, na.rm=T),
            max_age = max(age_months, na.rm=T))


# Age range for final month data entry
mean_age_final_entry_stringent <- clean_demog_stringent %>%
  filter(monolingual != 1) %>%
  arrange(subject_id, n_months) %>%
  # get the last row of data within each participant
  group_by(subject_id) %>%
  slice(n()) %>%
  ungroup() %>%
  # summarize
  summarize(mean_age = mean(age_months, na.rm=T),
            sd_age = sd(age_months, na.rm=T),
            min_age = min(age_months, na.rm=T),
            max_age = max(age_months, na.rm=T)) # the oldest age is 27m

# Total number of completed CDIs
## Completed English CDIs
n_eng_administration_stringent <- clean_demog_stringent  %>%
  filter(monolingual != 1) %>%
  select(c(subject_id, n_months, completed_eng)) %>%
  group_by(subject_id) %>%
  filter(!is.na(completed_eng)) %>%
  nrow()
  
## Completed French CDIs
n_fr_administration_stringent <- clean_demog_stringent  %>%
  filter(monolingual != 1) %>%
  select(c(subject_id, n_months, completed_fr)) %>%
  group_by(subject_id) %>%
  filter(!is.na(completed_fr)) %>%
  nrow()

## Completed both English and French CDIs
n_both_administration_stringent <- clean_demog_stringent %>%
  filter(monolingual != 1) %>%
  select(c(subject_id, n_months, completed_both)) %>%
  group_by(subject_id) %>%
  filter(completed_both == TRUE) %>%
  nrow()

## number of participants after retaining completed both English and French CDIs
n_subject_both_administration_stringent <- clean_demog_stringent %>%
  filter(monolingual != 1) %>%
  select(c(subject_id, n_months, completed_both)) %>%
  group_by(subject_id) %>%
  filter(completed_both == TRUE) %>%
  distinct(subject_id) %>%
  nrow()

# Number of participants who contributed data more than one time point
n_multiple_entry_stringent <- clean_demog_stringent %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  group_by(subject_id) %>%
  dplyr::count(subject_id) %>%
  filter(n > 1) %>%
  nrow()

# Average number of completed CDIs
n_avg_entries_stringent <- clean_demog_stringent  %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  group_by(subject_id) %>%
  dplyr::count(completed_eng) %>%
  filter(!is.na(completed_eng)) %>%
  ungroup()%>%
  summarize(mean_entry = mean(n, na.rm=T),
            sd_entry = sd(n, na.rm=T),
            min_entry = min(n, na.rm=T),
            max_entry = max(n, na.rm=T))
```

Using the more stringent  25%-75% language exposure criterion, we further eliminated a total of `r paste(exceed_25_75_langexp %>% nrow())` bilingual infants who had a wider language exposure range. This left us with data from `r paste(total_n_stringent)` participants (`r paste(n_female_stringent)` girls; mean starting age = `r round(mean_age_1st_entry_stringent$mean_age, 2)` months, $SD$ = `r round(mean_age_1st_entry_stringent$sd_age, 2)`, range = `r round(mean_age_1st_entry_stringent$min_age, 2)` – `r round(mean_age_1st_entry_stringent$max_age, 2)`; mean ending age = `r round(mean_age_final_entry_stringent$mean_age, 2)` months, $SD$ = `r round(mean_age_final_entry_stringent$sd_age, 2)`, range = `r round(mean_age_final_entry_stringent$min_age, 2)` – `r round(mean_age_final_entry_stringent$max_age, 2)`), with `r paste(n_eng_administration_stringent)` English CDI administrations and `r paste(n_fr_administration_stringent)` French CDI administration. As in the main analysis, we retained only those administrations where both the English and French were completed at the same time point; this gave us `r paste(n_both_administration_stringent)` completed administrations from `r paste(n_subject_both_administration_stringent)` infants. Among the `r paste(n_subject_both_administration_stringent)` infants, `r paste(n_subject_both_administration_stringent - n_multiple_entry_stringent)` infants contributed data at only one time point, and `r paste(n_multiple_entry_stringent)` infants contributed data at more than one time point, with participants contributing an average of `r round(n_avg_entries_stringent$mean_entry, 2)` measurements for each language ($SD$ = `r round(n_avg_entries_stringent$sd_entry, 2)`, range = `r round(n_avg_entries_stringent$min_entry)` – `r round(n_avg_entries_stringent$max_entry)`).

```{r include=FALSE, cache=TRUE}
# Average language exposure (weighted mean is calculated as participants contributed different number of LEQ update data)
## weighted mean across all participants
mean_langexp_stringent <- clean_demog_stringent %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  mutate(count = n()) %>%
  ungroup() %>%
  pivot_longer(c(exposure_eng, exposure_fr, exposure_other), names_to = "language", values_to = "exposure") %>%
  group_by(language) %>%
  summarize(mean_exposure = round(weightedMean(exposure, count), 1),
            sd_exposure = round(weightedSd(exposure, count), 1))

## minimum and maximum of individual averages
range_langexp_stringent <- clean_demog_stringent %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant
  summarize(mean_exposure_eng = mean(exposure_eng),
            mean_exposure_fr = mean(exposure_fr),
            mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  # minimum and maximum of the individual averages
  pivot_longer(c(mean_exposure_eng, mean_exposure_fr, mean_exposure_other), names_to = "language", values_to = "mean_exposure") %>%
  group_by(language) %>%
  summarize(min_exposure = min(mean_exposure),
            max_exposure = max(mean_exposure))

# Number of English dominant
## weighted mean across all participants
mean_engdom_stringent <- clean_demog_stringent %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant to define dominance
  mutate(count = n(),
         mean_exposure_eng = mean(exposure_eng),
         mean_exposure_fr = mean(exposure_fr),
         mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  filter(mean_exposure_eng > mean_exposure_fr) %>% # English exposure > French exposure
  select(-c(matches("mean"))) %>%
  # calculating weigthed mean
  pivot_longer(c(exposure_eng, exposure_fr, exposure_other), names_to = "language", values_to = "exposure") %>%
  group_by(language) %>%
  summarize(number = length(unique(subject_id)),
            mean_exposure = round(weightedMean(exposure, count), 1),
            sd_exposure = round(weightedSd(exposure, count), 1))

## minimum and maximum of individual averages
range_engdom_stringent <-clean_demog_stringent %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant
  summarize(mean_exposure_eng = mean(exposure_eng),
            mean_exposure_fr = mean(exposure_fr),
            mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  # English exposure > French exposure
  filter(mean_exposure_eng > mean_exposure_fr) %>%
  # minimum and maximum of the individual averages
  pivot_longer(c(mean_exposure_eng, mean_exposure_fr, mean_exposure_other), names_to = "language", values_to = "mean_exposure") %>%
  group_by(language) %>%
  summarize(number = length(unique(subject_id)),
            min_exposure = min(mean_exposure),
            max_exposure = max(mean_exposure))

# Number of French dominant
## weighted mean across all participants
mean_frdom_stringent <- clean_demog_stringent %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant to define dominance
  mutate(count = n(),
         mean_exposure_eng = mean(exposure_eng),
         mean_exposure_fr = mean(exposure_fr),
         mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  filter(mean_exposure_eng < mean_exposure_fr) %>% # English exposure < French exposure
  select(-c(matches("mean"))) %>%
  # calculating weigthed mean
  pivot_longer(c(exposure_eng, exposure_fr, exposure_other), names_to = "language", values_to = "exposure") %>%
  group_by(language) %>%
  summarize(number = length(unique(subject_id)),
            mean_exposure = round(weightedMean(exposure, count), 1),
            sd_exposure = round(weightedSd(exposure, count), 1))

## minimum and maximum of individual averages
range_frdom_stringent <- clean_demog_stringent %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant
  summarize(mean_exposure_eng = mean(exposure_eng),
            mean_exposure_fr = mean(exposure_fr),
            mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  # English exposure < French exposure
  filter(mean_exposure_eng < mean_exposure_fr) %>%
  # minimum and maximum of the individual averages
  pivot_longer(c(mean_exposure_eng, mean_exposure_fr, mean_exposure_other), names_to = "language", values_to = "mean_exposure") %>%
  group_by(language) %>%
  summarize(number = length(unique(subject_id)),
            min_exposure = min(mean_exposure),
            max_exposure = max(mean_exposure))

# Number of equal exposure to English and French
n_equalexp_stringent <- clean_demog_stringent %>%
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, n_months, exposure_eng, exposure_fr, exposure_other) %>%
  filter(!is.na(exposure_eng)) %>%
  group_by(subject_id) %>%
  # averaging within participant
  summarize(mean_exposure_eng = mean(exposure_eng),
            mean_exposure_fr = mean(exposure_fr),
            mean_exposure_other = mean(exposure_other)) %>%
  ungroup() %>%
  # English exposure = French exposure
  filter(mean_exposure_eng == mean_exposure_fr) %>%
  # count how many 
  nrow()

# Maternal education level in years
mean_maternaledu_stringent <- clean_demog_stringent %>% 
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, primary_caregiver_eng, primary_caregiver_education_eng) %>% # keeping "primary_caregiver_education_eng" because all listed mother as the primary caregiver
  distinct(subject_id, .keep_all = TRUE) %>%
  group_by(primary_caregiver_eng) %>%
  summarize(mean_edu_level = mean(primary_caregiver_education_eng, na.rm=T),
            sd_edu_level = sd(primary_caregiver_education_eng, na.rm=T),
            min_edu_level = min(primary_caregiver_education_eng, na.rm=T),
            max_edu_level = max(primary_caregiver_education_eng, na.rm=T))
  
# how many mothers completed a university degree or higher (years 16 = university graduate, > 16 = advanced degree)
n_maternal_uni_stringent <- clean_demog_stringent %>% 
  filter(monolingual != 1 & !is.na(completed_both)) %>%
  select(subject_id, primary_caregiver_eng, primary_caregiver_education_eng) %>% # keeping "primary_caregiver_education_eng" because all listed mother as the primary caregiver
  distinct(subject_id, .keep_all = TRUE) %>%
  # university graduate? (1 = yes, 0 = no)
  mutate(university_degree = if_else(primary_caregiver_education_eng >= 16, 1, 0)) %>%
  # summary
  dplyr::count(university_degree) %>%
  summarize(university_degree = university_degree,
            n = n, 
            percentage = n/sum(n)*100)

```

On average, this group of `r  paste(n_subject_both_administration_stringent)` infants were exposed to English `r round(mean_langexp_stringent$mean_exposure[mean_langexp_stringent$language=="exposure_eng"],1)`% of the time ($SD$ = `r round(mean_langexp_stringent$sd_exposure[mean_langexp_stringent$language=="exposure_eng"],1)`, range = `r range_langexp_stringent$min_exposure[range_langexp_stringent$language=="mean_exposure_eng"]` – `r range_langexp_stringent$max_exposure[range_langexp_stringent$language=="mean_exposure_eng"]`), to French `r round(mean_langexp_stringent$mean_exposure[mean_langexp_stringent$language=="exposure_fr"],1)`% of the time ($SD$ = `r round(mean_langexp_stringent$sd_exposure[mean_langexp_stringent$language=="exposure_fr"],1)`, range = `r range_langexp_stringent$min_exposure[range_langexp_stringent$language=="mean_exposure_fr"]` – `r range_langexp_stringent$max_exposure[range_langexp_stringent$language=="mean_exposure_fr"]`), and to a third language `r round(mean_langexp_stringent$mean_exposure[mean_langexp_stringent$language=="exposure_other"],1)`% of the time (SD = `r round(mean_langexp_stringent$sd_exposure[mean_langexp_stringent$language=="exposure_other"],1)`, range = `r range_langexp_stringent$min_exposure[range_langexp_stringent$language=="mean_exposure_other"]` – `r range_langexp_stringent$max_exposure[range_langexp_stringent$language=="mean_exposure_other"]`). `r paste(mean_engdom_stringent$number[1])` of the bilingual infants were English-dominant ($M$ = `r round(mean_engdom_stringent$mean_exposure[mean_engdom_stringent$language=="exposure_eng"],1)`% English exposure, $SD$ = `r round(mean_engdom_stringent$sd_exposure[mean_engdom_stringent$language=="exposure_eng"],1)`, range = `r range_engdom_stringent$min_exposure[range_engdom_stringent$language=="mean_exposure_eng"]` – `r range_engdom_stringent$max_exposure[range_engdom_stringent$language=="mean_exposure_eng"]`), `r paste(mean_frdom_stringent$number[1])` were French-dominant  ($M$ = `r round(mean_frdom_stringent$mean_exposure[mean_frdom_stringent$language=="exposure_fr"],1)`% French exposure, $SD$ = `r round(mean_frdom_stringent$sd_exposure[mean_frdom_stringent$language=="exposure_fr"],1)`, range = `r range_frdom_stringent$min_exposure[range_frdom_stringent$language=="mean_exposure_fr"]` – `r range_frdom_stringent$max_exposure[range_frdom_stringent$language=="mean_exposure_fr"]`), and `r paste(n_equalexp_stringent)` reported equal exposure to both English and French. The average maternal education level was `r round(mean_maternaledu_stringent$mean_edu_level,2)` years ($SD$ = `r round(mean_maternaledu_stringent$sd_edu_level,2)`, range = `r round(mean_maternaledu_stringent$min_edu_level)` – `r round(mean_maternaledu_stringent$max_edu_level)`), and `r round(n_maternal_uni_stringent$percentage[n_maternal_uni_stringent$university_degree=="1"],2)`% of the mothers had completed a university degree or higher. 

## Descriptive Measures of Number of Words Produced
```{r include=FALSE, cache=TRUE}
# complete list
keepers_cognate_full <- read.csv(here::here("data_keepers/public_keepers_cognate_full.csv"))

## calculate to see: (1) how many total words were produced; (2) if the TE pairs have been produced (i.e., both words in English and French are acquired); (3) how many cognate or non-cognate words have been produced (i.e., either word acquired only in one language or words acquired in both language)
keepers_cognate_full_wide <- keepers_cognate_full %>%
  # remove unnecessary variables
  select(-c(english_item, french_item)) %>%
  # calculate total number of eng/fr words acquired
  group_by(subject_id, n_months) %>%
  mutate(total_eng_acquired = sum(eng_acquired, na.rm=T),
         total_fr_acquired = sum(fr_acquired, na.rm=T)) %>%
  ungroup() %>%
  # calculate total number of words acquired (eng + fr)
  mutate(total_acquired = total_eng_acquired + total_fr_acquired,
         # percentage of words acquired
         percent_word_acquired = total_acquired/(537*2)*100) %>%
  # calculate to see if the TE pairs have been produced
  mutate(TE_acquired = case_when(eng_acquired == 1 & fr_acquired == 1 ~ 1,
                                 TRUE ~ 0)) %>%
  # remove the eng_acquired and fr_acquired variables
  select(-c(eng_acquired, fr_acquired)) %>%
  # unite french_item_id with word_pairs to facilitate the pivot_wider step
  unite(word_pairs_withID, c(english_item_id, word_pairs, french_item_id), remove=FALSE) %>%
  select(-c(word_pairs, english_item_id, french_item_id)) %>%
  # create wide data
  pivot_wider(names_from = word_pairs_withID, values_from = TE_acquired) %>%
  mutate(total_TE_acquired = rowSums(.[17:553], na.rm=T),
         percent_TE_acquired = case_when(cognate_status == "cognate" ~ total_TE_acquired/131*100,
                                       cognate_status == "non-cognate" ~ total_TE_acquired/406*100)) %>%
  # calculate total number of TEs (regardless of cognate status)
  group_by(subject_id, n_months) %>%
  mutate(total_all_TE_acquired = sum(total_TE_acquired, na.rm=T),
         percent_all_TE_acquired = total_all_TE_acquired/537*100) %>%
  ungroup()

# matched list
keepers_cognate_matched <- read.csv(here::here("data_keepers/public_keepers_cognate_matched.csv"))

## calculate to see: (1) how many total words were produced; (2) if the TE pairs have been produced (i.e., both words in English and French are acquired)
## (3) how many cognate or non-cognate words have been produced (i.e., either word acquired only in one language or words acquired in both language)
keepers_cognate_matched_wide <- keepers_cognate_matched %>%
  # remove unnecessary variables
  select(-c(english_item_id, english_item, Eng_AoA, french_item_id, french_item, Fr_AoA)) %>%
  # calculate total number of eng/fr words acquired within the matched list
  group_by(subject_id, n_months) %>%
  mutate(total_eng_acquired = sum(eng_acquired, na.rm=T),
         total_fr_acquired = sum(fr_acquired, na.rm=T)) %>%
  ungroup() %>%
  # calculate total number of words acquired (eng + fr)
  mutate(total_acquired = total_eng_acquired + total_fr_acquired,
         # percentage of words acquired
         percent_word_acquired = total_acquired/(162*2)*100) %>%
  # calculate to see if the TE pairs have been produced
  mutate(TE_acquired = case_when(eng_acquired == 1 & fr_acquired == 1 ~ 1,
                                 TRUE ~ 0))  %>%
  # remove the eng_acquired and fr_acquired variables
  select(-c(eng_acquired, fr_acquired)) %>%
  pivot_wider(names_from = word_pairs, values_from = TE_acquired) %>%
  mutate(total_TE_acquired = rowSums(.[17:178], na.rm=T)) %>%
  mutate(percent_TE_acquired = total_TE_acquired/81*100) %>%
  # calculate total number of TEs (regardless of cognate status)
  group_by(subject_id, n_months) %>%
  mutate(total_all_TE_acquired = sum(total_TE_acquired, na.rm=T),
         percent_all_TE_acquired = total_all_TE_acquired/162*100) %>%
  ungroup()
```

```{r include=FALSE, cache=TRUE}
# ---------- 25-75% language exposure criterion: Analysis 1 ----------

keepers_cognate_full_stringent <- keepers_cognate_full %>%
  anti_join(exceed_25_75_langexp)

keepers_cognate_matched_stringent <- keepers_cognate_matched %>%
  anti_join(exceed_25_75_langexp)
  
keepers_cognate_full_wide_stringent <- keepers_cognate_full_wide %>%
  anti_join(exceed_25_75_langexp)

keepers_cognate_matched_wide_stringent <- keepers_cognate_matched_wide %>%
  anti_join(exceed_25_75_langexp)


# descriptives
## mean total words acquired
full_mean_total_stringent <- keepers_cognate_full_wide_stringent %>%
  select(subject_id, n_months, total_eng_acquired, total_fr_acquired, total_acquired, percent_word_acquired, 
         total_all_TE_acquired, percent_all_TE_acquired) %>%
  distinct(subject_id, n_months, .keep_all = TRUE) %>%
  pivot_longer(-c(subject_id, n_months), names_to = "type", values_to = "number") %>%
  group_by(type) %>%
  summarize(n_subject = n_distinct(subject_id),
            mean_number = mean(number, na.rm=T),
            sd_number = sd(number, na.rm=T),
            min_number = min(number, na.rm=T),
            max_number = max(number, na.rm=T))

matched_mean_total_stringent <- keepers_cognate_matched_wide_stringent %>%
  select(subject_id, n_months, total_eng_acquired, total_fr_acquired, total_acquired, percent_word_acquired, 
         total_all_TE_acquired, percent_all_TE_acquired) %>%
  distinct(subject_id, n_months, .keep_all = TRUE) %>%
  pivot_longer(-c(subject_id, n_months), names_to = "type", values_to = "number") %>%
  group_by(type) %>%
  summarize(n_subject = n_distinct(subject_id),
            mean_number = mean(number, na.rm=T),
            sd_number = sd(number, na.rm=T),
            min_number = min(number, na.rm=T),
            max_number = max(number, na.rm=T))

# Dependent Variable 1: Cognate Words Versus Non-Cognate Words
## average age in the sample
mean_age_stringent <- keepers_cognate_full_stringent %>% 
  distinct(subject_id, .keep_all = T) %>% 
  summarize(mean_age = mean(age_days, na.rm = T))

## Complete List
### calculate to see how many cognate or non-cognate words have been produced (i.e., either word acquired only in one language or words acquired in both language)
keepers_cognate_full_notPair_stringent <- keepers_cognate_full_stringent %>%
  # remove unnecessary variables
  select(-c(english_item_id, english_item, french_item_id, french_item)) %>%
  # calculate total number of eng/fr words acquired within the full list
  group_by(subject_id, n_months, age_days, cognate_status) %>%
  summarize(n_Eng_acquired = sum(eng_acquired, na.rm=T),
            n_Fr_acquired = sum(fr_acquired, na.rm=T))%>%
  # calculate total number of words acquired within the full list
  mutate(n_word_acquired = n_Eng_acquired + n_Fr_acquired,
         # percentage of words acquired within the full list
         percent_word_acquired = case_when(cognate_status == "cognate" ~ n_word_acquired/(131*2)*100,
                                       cognate_status == "non-cognate" ~ n_word_acquired/(406*2)*100),
         # proportion of words acquired within the full list
         prop_word_acquired = case_when(cognate_status == "cognate" ~ n_word_acquired/(131*2),
                                       cognate_status == "non-cognate" ~ n_word_acquired/(406*2))) %>%
  ungroup() 

### descriptive: mean number of cognates/non-cognates acquired
full_mean_cognates_stringent <- keepers_cognate_full_notPair_stringent %>%
  group_by(cognate_status) %>%
  summarize(mean_n_word = mean(n_word_acquired, na.rm=T),
            sd_n_word = sd(n_word_acquired, na.rm=T),
            min_n_word = min(n_word_acquired, na.rm=T),
            max_n_word = max(n_word_acquired, na.rm=T))

full_mean_cognates_prop_stringent <- keepers_cognate_full_notPair_stringent %>%
  group_by(cognate_status) %>%
  summarize(mean_prop_word = mean(prop_word_acquired, na.rm=T),
            sd_prop_word = sd(prop_word_acquired, na.rm=T),
            min_prop_word = min(prop_word_acquired, na.rm=T),
            max_prop_word = max(prop_word_acquired, na.rm=T))

### logistic mixed-effects model
stringent_Analysis1_full_logistic_model <- keepers_cognate_full_notPair_stringent %>%
  # scale and center age for ease of interpretation
  mutate(age_days = scale(age_days, scale = FALSE, center = TRUE)) %>%
  # reverse code "cognate_status"
  mutate(cognate_status = fct_rev(cognate_status)) %>%
  # create a weight variable for the model
  mutate(total_number = case_when(cognate_status == "cognate" ~ 131*2,
                        cognate_status == "non-cognate" ~ 406*2)) %>%
  # the glmer model
  glmer(prop_word_acquired ~ cognate_status * age_days + (1|subject_id),
        family = binomial, weights = total_number,
        data = .,
        nAGQ = 0)

stringent_Analysis1_full_logistic_model_coefs <- summary(stringent_Analysis1_full_logistic_model)$coef %>%
  as.data.frame %>%
  mutate_at(c("Estimate","Std. Error", "z value", "Pr(>|z|)"), 
            function (x) signif(x, digits = 3)) %>% 
  rename(SE = `Std. Error`, 
         z = `z value`,
         p = `Pr(>|z|)`) %>%
  mutate_if(is.numeric, round, digits = 3) %>%
  mutate(p = case_when(
    p < .001 ~ "<.001",
    p < .01 ~ "<.01",
    p < .05 ~ "<.05",
    TRUE ~ as.character(p)))

rownames(stringent_Analysis1_full_logistic_model_coefs) <- c("Intercept", "cognate_status", "age_days", 
                     "cognate_status * age_days") 

## Matched List
### calculate to see how many cognate or non-cognate words have been produced (i.e., either word acquired only in one language or words acquired in both language)
keepers_cognate_matched_notPair_stringent <- keepers_cognate_matched_stringent %>%
  # remove unnecessary variables
  select(-c(english_item_id, english_item, Eng_AoA, french_item_id, french_item, Fr_AoA)) %>%
  # calculate total number of eng/fr words acquired within the matched list
  group_by(subject_id, n_months, age_days, cognate_status) %>%
  summarize(n_Eng_acquired = sum(eng_acquired, na.rm=T),
            n_Fr_acquired = sum(fr_acquired, na.rm=T))%>%
  # calculate total number of words acquired within the matched list
  mutate(n_word_acquired = n_Eng_acquired + n_Fr_acquired,
         percent_word_acquired = n_word_acquired/(81*2)*100, # percentage of words acquired within the matched list
         prop_word_acquired = n_word_acquired/(81*2)) %>%
  ungroup() 

### descriptive: mean number of cognates/non-cognates acquired
matched_mean_cognates_stringent <- keepers_cognate_matched_notPair_stringent %>%
  group_by(cognate_status) %>%
  summarize(mean_n_word = mean(n_word_acquired, na.rm=T),
            sd_n_word = sd(n_word_acquired, na.rm=T),
            min_n_word = min(n_word_acquired, na.rm=T),
            max_n_word = max(n_word_acquired, na.rm=T))

matched_mean_cognates_prop_stringent <- keepers_cognate_matched_notPair_stringent %>%
  group_by(cognate_status) %>%
  summarize(mean_prop_word = mean(prop_word_acquired, na.rm=T),
            sd_prop_word = sd(prop_word_acquired, na.rm=T),
            min_prop_word = min(prop_word_acquired, na.rm=T),
            max_prop_word = max(prop_word_acquired, na.rm=T))

### logistic mixed-effects model
stringent_Analysis1_matched_logistic_model <- keepers_cognate_matched_notPair_stringent %>%
  # scale and center age for ease of interpretation
  mutate(age_days = scale(age_days, scale = FALSE, center = TRUE)) %>%
  # reverse code "cognate_status"
  mutate(cognate_status = fct_rev(cognate_status)) %>%
  # create a weight variable for the model
  mutate(total_number = 81*2) %>%
  # the glmer model
  glmer(prop_word_acquired ~ cognate_status * age_days + (1|subject_id),
        family = binomial, weights = total_number,
        data = .,
        nAGQ = 0)

stringent_Analysis1_matched_logistic_model_coefs <- summary(stringent_Analysis1_matched_logistic_model)$coef %>%
  as.data.frame %>%
  mutate_at(c("Estimate","Std. Error", "z value", "Pr(>|z|)"), 
            function (x) signif(x, digits = 3)) %>% 
  rename(SE = `Std. Error`, 
         z = `z value`,
         p = `Pr(>|z|)`) %>%
  mutate_if(is.numeric, round, digits = 3) %>%
  mutate(p = case_when(
    p < .001 ~ "<.001",
    p < .01 ~ "<.01",
    p < .05 ~ "<.05",
    TRUE ~ as.character(p)))

rownames(stringent_Analysis1_matched_logistic_model_coefs) <- c("Intercept", "cognate_status", "age_days", 
                     "cognate_status * age_days") 

```

```{r FigS1, fig.cap="Proportion of words produced by age and cognate status, with Panel A representing the complete list and Panel B representing the matched list. Note that the black dashed line represents the mean age of 547.6 days which serves as the reference level for age in our models.", echo=FALSE, dpi=600, fig.align='center', fig.height=4, out.width="120%", cache = TRUE}

prop_word_full_stringent <- keepers_cognate_full_notPair_stringent %>%
  select(subject_id, n_months, age_days, cognate_status, prop_word_acquired) %>%
  rename(full_prop_word = prop_word_acquired) 

prop_word_matched_stringent <- keepers_cognate_matched_notPair_stringent %>%
  select(subject_id, n_months, age_days, cognate_status, prop_word_acquired) %>%
  rename(matched_prop_word = prop_word_acquired)

stringent_Figure1 <- right_join(prop_word_full_stringent, prop_word_matched_stringent) %>%
  pivot_longer(c(full_prop_word, matched_prop_word), names_to = "list", values_to = "prop_word") %>%
  mutate(list = recode(list, 
                        full_prop_word = "Panel A: Complete list",
                        matched_prop_word = "Panel B: Matched list"),
         cognate_status = recode(cognate_status,
                                  "cognate" = "Cognate",
                                  "non-cognate" = "Non-cognate")) %>%
  group_by(subject_id, n_months, list, cognate_status) %>%
  ggplot(aes(x = age_days, y = prop_word, color = cognate_status)) +
  facet_wrap(. ~ list) +
  geom_smooth(method = "glm", se = FALSE, method.args = list(family = binomial)) + 
  geom_point(stat = "identity", position = "identity", shape = 23, alpha = 0.3) +
  geom_vline(xintercept = 548.3, linetype="dashed", alpha = 0.5) +
  labs(x = "\n Age (in days)", 
       y = "Proportion of words produced",
       color = "Cognate status") + 
  ylim(0, 1) +
  theme_minimal() +
  theme(text = element_text(size=12),
        strip.text.x = element_text(size = 12),
        legend.position = "bottom",
        panel.spacing = unit(1.5, "lines")) 

stringent_Figure1

ggsave("stringent_Figure1.png", stringent_Figure1,
       width = 9, height = 6)

```

```{r Table S1, results = "asis", cache = TRUE}
TableS1 <- cbind(stringent_Analysis1_full_logistic_model_coefs, stringent_Analysis1_matched_logistic_model_coefs)

papaja::apa_table(TableS1, format.args = list(digits = 3),
                  col.names =c("","Estimate", "$SE$","$z$","$p$","Estimate", "$SE$","$z$","$p$"),
                  align=c("l","c","c","c","c","c","c","c", "c"),
                  col_spanners = list("Complete list" = c(2,5), "Matched list" = c(6,9)),
                  caption="Table S1. Coefficient estimates from the mixed-effects logistic models predicting proportion of words produced.",
                  landscape = FALSE)
```

```{r include=FALSE, cache=TRUE}
# ---------- 25-75% language exposure criterion: Analysis 2 ----------

# Dependent Variable 2: Cognate Pairs Versus Non-Cognate Pairs

## Complete List
### mean number of cognates/non-cognates acquired
#### total number
full_mean_total_TE_stringent <- keepers_cognate_full_wide_stringent %>%
  group_by(cognate_status) %>%
  summarize(mean_TE = mean(total_TE_acquired, na.rm=T),
            sd_TE = sd(total_TE_acquired, na.rm=T),
            min_TE = min(total_TE_acquired, na.rm=T),
            max_TE = max(total_TE_acquired, na.rm=T))

#### proportion
full_mean_prop_TE_stringent <- keepers_cognate_full_wide_stringent %>%
  mutate(prop_TE_acquired = percent_TE_acquired/100) %>%
  group_by(cognate_status) %>%
  summarize(mean_TE = mean(prop_TE_acquired, na.rm=T),
            sd_TE = sd(prop_TE_acquired, na.rm=T),
            min_TE = min(prop_TE_acquired, na.rm=T),
            max_TE = max(prop_TE_acquired, na.rm=T))

### logistic model
stringent_Analysis2_full_logistic_model <- keepers_cognate_full_wide_stringent %>%
  # reverse-order cognate status, so that the non-cognate status would be the reference level
  mutate(cognate_status = fct_rev(cognate_status)) %>%
  # scale and center age for ease of interpretation
  mutate(age_days = scale(age_days, scale = FALSE, center = TRUE)) %>%
  # change percentage to proportion
  mutate(prop_TE_acquired = percent_TE_acquired/100) %>%
  # create a weight variable for the model
  mutate(total_number = case_when(cognate_status == "cognate" ~ 131,
                        cognate_status == "non-cognate" ~ 406)) %>%
  # the glmer model
  glmer(prop_TE_acquired ~ cognate_status * age_days + (1|subject_id),
        family = binomial, weights = total_number,
        data = .,
        nAGQ = 0)

stringent_Analysis2_full_logistic_model_coefs <- summary(stringent_Analysis2_full_logistic_model)$coef %>%
  as.data.frame %>%
  mutate_at(c("Estimate","Std. Error", "z value", "Pr(>|z|)"), 
            function (x) signif(x, digits = 3)) %>% 
  rename(SE = `Std. Error`, 
         z = `z value`,
         p = `Pr(>|z|)`) %>%
  mutate_if(is.numeric, round, digits = 3) %>%
  mutate(p = case_when(
    p < .001 ~ "<.001",
    p < .01 ~ "<.01",
    p < .05 ~ "<.05",
    TRUE ~ as.character(p)))

rownames(stringent_Analysis2_full_logistic_model_coefs) <- c("Intercept", "cognate_status", "age_days", 
                     "cognate_status * age_days") 


## Matched List
### mean number of cognates/non-cognates acquired
#### total number
matched_mean_total_TE_stringent <- keepers_cognate_matched_wide_stringent %>%
  group_by(cognate_status) %>%
  summarize(mean_TE = mean(total_TE_acquired, na.rm=T),
            sd_TE = sd(total_TE_acquired, na.rm=T),
            min_TE = min(total_TE_acquired, na.rm=T),
            max_TE = max(total_TE_acquired, na.rm=T))

#### proportion
matched_mean_prop_TE_stringent <- keepers_cognate_matched_wide_stringent %>%
  mutate(prop_TE_acquired = percent_TE_acquired/100) %>%
  group_by(cognate_status) %>%
  summarize(mean_TE = mean(prop_TE_acquired, na.rm=T),
            sd_TE = sd(prop_TE_acquired, na.rm=T),
            min_TE = min(prop_TE_acquired, na.rm=T),
            max_TE = max(prop_TE_acquired, na.rm=T))

### logistic model
stringent_Analysis2_matched_logistic_model <- keepers_cognate_matched_wide_stringent %>%
  # reverse-order cognate status, so that the non-cognate status would be the reference level
  mutate(cognate_status = fct_rev(cognate_status)) %>%
  # scale and center age for ease of interpretation
  mutate(age_days = scale(age_days, scale = FALSE, center = TRUE)) %>%
  # change percentage to proportion
  mutate(prop_TE_acquired = percent_TE_acquired/100) %>%
  # create a weight variable for the model
  mutate(total_number = 81) %>%
  # the glmer model
  glmer(prop_TE_acquired ~ cognate_status * age_days + (1|subject_id),
        family = binomial, weights = total_number,
        data = .,
        nAGQ = 0)

stringent_Analysis2_matched_logistic_model_coefs <- summary(stringent_Analysis2_matched_logistic_model)$coef %>%
  as.data.frame %>%
  mutate_at(c("Estimate","Std. Error", "z value", "Pr(>|z|)"), 
            function (x) signif(x, digits = 3)) %>% 
  rename(SE = `Std. Error`, 
         z = `z value`,
         p = `Pr(>|z|)`) %>%
  mutate_if(is.numeric, round, digits = 3) %>%
  mutate(p = case_when(
    p < .001 ~ "<.001",
    p < .01 ~ "<.01",
    p < .05 ~ "<.05",
    TRUE ~ as.character(p)))

rownames(stringent_Analysis2_matched_logistic_model_coefs) <- c("Intercept", "cognate_status", "age_days", 
                     "cognate_status * age_days") 

```

```{r FigS2, fig.cap="Proportion of translation equivalent pairs produced by age and cognate status, with Panel A representing the complete list and Panel B representing the matched list. Note that the black dashed line represents the mean age of 547.6 days which serves as the reference level for age in our models.", echo=FALSE, dpi=600, fig.align='center', fig.height=4, out.width="120%", cache = TRUE}

prop_TE_full_stringent <- keepers_cognate_full_wide_stringent %>%
  select(subject_id, n_months, age_days, cognate_status, percent_TE_acquired) %>%
  mutate(full_prop_TE = percent_TE_acquired/100) %>%
  select(-(percent_TE_acquired))

prop_TE_matched_stringent <- keepers_cognate_matched_wide_stringent %>%
  select(subject_id, n_months, age_days, cognate_status, percent_TE_acquired) %>%
  mutate(matched_prop_TE = percent_TE_acquired/100) %>%
  select(-(percent_TE_acquired))

stringent_Figure2 <- right_join(prop_TE_full_stringent, prop_TE_matched_stringent) %>%
  pivot_longer(c(full_prop_TE, matched_prop_TE), names_to = "list", values_to = "prop_TE") %>%
  mutate(list = recode(list, 
                        full_prop_TE = "Panel A: Complete list",
                        matched_prop_TE = "Panel B: Matched list"),
         cognate_status = recode(cognate_status,
                                  "cognate" = "Cognate",
                                  "non-cognate" = "Non-cognate")) %>%
  group_by(subject_id, n_months, list, cognate_status) %>%
  ggplot(aes(x = age_days, y = prop_TE, color = cognate_status)) +
  facet_wrap(. ~ list) +
  geom_smooth(method = "glm", se = FALSE, method.args = list(family = binomial)) + 
  geom_point(stat = "identity", position = "identity", shape = 23, alpha = 0.3) +
  geom_vline(xintercept = 548.3, linetype = "dashed", alpha = 0.5) +
  labs(x = "\n Age (in days)", 
       y = "Proportion of translation equivalent pairs produced",
       color = "Cognate status") + 
  ylim(0,1) +
  theme_minimal() +
  theme(text = element_text(size=12),
        strip.text.x = element_text(size = 12),
        legend.position = "bottom",
        panel.spacing = unit(1.5, "lines")) 

stringent_Figure2

ggsave("stringent_Figure2.png", stringent_Figure2,
       width = 9, height = 6)
```

## Summary of Analyses
Overall, the result patterns observed here with the more stringent language exposure criterion were consistent with those reported in the main analysis. This suggests that the cognate effect we observed was robust across different language exposure inclusion criteria.
